---
title: "LUCAS dataset preparation for the OSSL"
author: 
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Ran Zhi
    email: jsafanelli@woodwellclimate.org
  - name: Tomislav Hengl
    email: tom.hengl@opengeohub.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Brazilian Soil Spectral
Library (lucas) into the Open Soil Spectral Library.

Website: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The LUCAS data

Soil samples from the LUCAS (Land Use/Cover Area frame statistical Survey) of 23-28 European Union member states collected between 2009 and 2015 were scanned for VisNIR (@Stevens2013, @Nocita2014).

The LUCAS topsoil data (@Orgiazzi2017) used in this work was made available by the European Commission through the European Soil Data Centre managed by the [Joint Research Centre (JRC)](http://esdac.jrc.ec.europa.eu/).

JRC has a restrictive license for redistribution, but permission was granted to share it through the OSSL under the condition that “Graphical representation of individual units on a map is permitted as far as the geographical location of the soil samples is not detectable”.

Additional 600 samples from the LUCAS soil samples archive were scanned at Woodwell Climate Research Center using a MIR instrument @Dangal2020.

Input datasets:

- `LUCAS.SOIL_corr.Rdata`: VNIR spectra.
- `LUCAS_Topsoil_2009_ESPG4326.csv`: Site and soil from 2009.
- `LUCAS_spectra_2015.rds`: VNIR from 2015.
- `LUCAS_Topsoil_complete_2015_ESPG4326.csv`: Site and soil from 2015.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "data.table", "lubridate", "units", "readxl", "stringr",
              "googledrive", "googlesheets4", "tmap", "sf", "rnaturalearth",
              "skimr", "prospectr", "tictoc", "nanoparquet", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
gs4_auth(path = "~/secrets/soilcarbon-soilspec-845d0d3f2fbe.json")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r dir_ref}
dir = "~/mnt-ossl-private/database/datasets/LUCAS"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r soilsite}
# Data from 2009-2012 survey
gpkg.lst <- list.files(dir, glob2rx("SoilAttr_*.gpkg$"), full.names = TRUE)

lucas.2009 <- lapply(gpkg.lst, function(i){
  sf::st_read(i) %>%
    dplyr::mutate(lon = sf::st_coordinates(.)[,1],
                  lat = sf::st_coordinates(.)[,2]) %>%
    dplyr::as_tibble(.) %>%
    dplyr::select(-geom)})

lucas.2009 <- Reduce(dplyr::bind_rows, lucas.2009) %>%
  select(POINT_ID, lon, lat, Country) %>%
  mutate(POINT_ID = paste0("2009.", POINT_ID)) %>%
  rename(id.layer_local_c = POINT_ID,
         longitude.point_wgs84_dd = lon,
         latitude.point_wgs84_dd = lat,
         loc.country_src_txt = Country) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  mutate(layer.upper.depth_usda_cm = 0,
         layer.lower.depth_usda_cm = 20,
         observation.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2009-05-01"),
         observation.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2012-08-01"))

# Data from 2015 survey
lucas.2015.x1 <- fread(path(dir, "/LUCAS_Topsoil_complete_2015_ESPG4326.csv"))

# lucas.2015 <- sf::st_read(path(dir, "/LUCAS_2015.gpkg")) %>%
#   dplyr::mutate(lon = sf::st_coordinates(.)[,1],
#                 lat = sf::st_coordinates(.)[,2]) %>%
#   dplyr::as_tibble(.) %>%
#   dplyr::select(-geom)

lucas.2015 <- lucas.2015.x1 %>%
  select(Point_ID, xcoord, ycoord, ADMIN, LC0_Desc) %>%
  mutate(Point_ID = paste0("2015.", Point_ID)) %>%
  rename(id.layer_local_c = Point_ID,
         longitude.point_wgs84_dd = xcoord,
         latitude.point_wgs84_dd = ycoord,
         loc.country_src_txt = ADMIN,
         site.landuse_src_txt = LC0_Desc) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  mutate(layer.upper.depth_usda_cm = 0,
         layer.lower.depth_usda_cm = 20,
         observation.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2015-03-01"),
         observation.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2015-10-01"))

# Binding datasets
# Degrade location accuracy +/-1000m so we can share the points
lucas.sitedata <- bind_rows(lucas.2009, lucas.2015) %>% 
  mutate(longitude.point_wgs84_dd = longitude.point_wgs84_dd + rnorm(n(), 0, 0.005),
         latitude.point_wgs84_dd = latitude.point_wgs84_dd + rnorm(n(), 0, 0.005)) %>% 
  # mutate(id.project_ascii_txt = "Land Use and Coverage Area frame Survey (LUCAS) topsoil data",
  #        id.dataset.site_ascii_txt = str_sub(id.layer_local_c, 6, -1),
  #        layer.texture_usda_txt = "",
  #        pedon.taxa_usda_txt = "",
  #        horizon.designation_usda_txt = "",
  #        longitude.county_wgs84_dd = NA,
  #        latitude.county_wgs84_dd = NA,
  #        location.point.error_any_m = 1000,
  #        location.country_iso.3166_txt = "",
  #        observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
  #        observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
  #        surveyor.title_utf8_txt = "Institute for Environment and Sustainability (Joint Research Centre)",
  #        surveyor.contact_ietf_email = "ec-esdac@jrc.ec.europa.eu",
  #        surveyor.address_utf8_txt = "",
  #        dataset.title_utf8_txt = "LUCAS 2009, 2015 topsoil data",
  #        dataset.owner_utf8_txt = "European Soil Data Centre (ESDAC), European Commission, Joint Research Centre",
  #        dataset.code_ascii_txt = "LUCAS.SSL",
  #        dataset.address_idn_url = "https://esdac.jrc.ec.europa.eu/resource-type/soil-point-data",
  #        dataset.license.title_ascii_txt = "JRC License Agreement",
  #        dataset.license.address_idn_url = "https://esdac.jrc.ec.europa.eu/resource-type/soil-point-data",
  #        dataset.contact.name_utf8_txt = "ESDAC - European Commissiony",
  #        dataset.contact_ietf_email = "ec-esdac@jrc.ec.europa.eu") %>%
  mutate(dataset.code_ascii_txt = "LUCAS",
         id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .before = 1)

# Removing duplicates
lucas.sitedata %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  group_by(repeats) %>%
  summarise(count = n())

dupli.ids <- lucas.sitedata %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  filter(repeats > 1) %>%
  pull(id.layer_local_c)

lucas.sitedata <- lucas.sitedata %>%
  filter(!(id.layer_local_c %in% dupli.ids)) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.exp.file = path(dir, "ossl_soilsite_v1.3")
readr::write_csv(lucas.sitedata, str_c(site.exp.file, ".csv.gz"))
nanoparquet::write_parquet(lucas.sitedata, str_c(site.exp.file, ".parquet"))
```

Plotting sites map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

ocean <- ne_download(scale = 110, type = "ocean", category = "physical", returnclass = "sf")

points <- lucas.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")

tm_shape(ocean) +
  tm_polygons(fill = "lightblue", col = NA) +
  tm_shape(World) +
  tm_polygons('#f0f0f0f0', col_alpha = 0.2) +
  tm_shape(points) +
  tm_dots(size = 0.10, fill = "firebrick") +
  tm_crs("ESRI:54030") +
  tm_layout(frame = FALSE)
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwriting, then rename online to download locally.**

```{r soil_template, echo=TRUE, eval=FALSE}
gpkg.lst = list.files(dir, glob2rx("SoilAttr_*.gpkg$"), full.names = TRUE)
lucas.2009 = lapply(gpkg.lst, function(i){sf::st_read(i) %>% dplyr::as_tibble(.) %>% dplyr::select(-geom)})
lucas.2009 = Reduce(dplyr::bind_rows, lucas.2009)

lucas.2015 = fread(paste0(dir, "/LUCAS_Topsoil_2015_20200323.csv"))

soillab.names <- lucas.2009 %>%
  names(.) %>%
  tibble::tibble(original_name = .) %>%
  dplyr::mutate(table = 'LUCAS 2009/2012; SoilAttr_*.gpkg', .before = 1) %>%
  dplyr::bind_rows({
    lucas.2015 %>%
      names(.) %>%
      tibble::tibble(original_name = .) %>%
      dplyr::mutate(table = 'LUCAS_Topsoil_2015.csv', .before = 1)
  }) %>%
  dplyr::mutate(import = '', ossl_name = '', .after = original_name) %>%
  dplyr::mutate(comment = '')

readr::write_csv(soillab.names, paste0(getwd(), "/lucas_soillab_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the edited standardization template is saved to this dataset folder.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "LUCAS")

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F) %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
# Data from 2009-2012 survey
gpkg.lst <- list.files(dir, glob2rx("SoilAttr_*.gpkg$"), full.names = TRUE)

lucas.2009 <- lapply(gpkg.lst, function(i){
  sf::st_read(i) %>%
    dplyr::as_tibble(.) %>%
    dplyr::select(-geom)})

lucas.2009 <- Reduce(dplyr::bind_rows, lucas.2009) %>%
  mutate(POINT_ID = paste0("2009.", POINT_ID))

# Harmonization of names and units
# 2009 analyte selection and names
analytes.2009.old.names <- transvalues %>%
  filter(table == "LUCAS 2009/2012; SoilAttr_*.gpkg") %>%
  filter(original_name != "POINT_ID") %>%
  pull(original_name)

analytes.2009.new.names <- transvalues %>%
  filter(table == "LUCAS 2009/2012; SoilAttr_*.gpkg") %>%
  filter(original_name != "POINT_ID") %>%
  pull(ossl_name)

# Selecting and renaming
lucas.2009 <- lucas.2009 %>%
  rename(id.layer_local_c = POINT_ID) %>%
  select(id.layer_local_c, all_of(analytes.2009.old.names)) %>%
  rename_with(~analytes.2009.new.names, all_of(analytes.2009.old.names))

# Removing duplicates
# lucas.2009 %>%
#   group_by(id.layer_local_c) %>%
#   summarise(repeats = n()) %>%
#   group_by(repeats) %>%
#   summarise(count = n())

dupli.ids.2009 <- lucas.2009 %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  filter(repeats > 1) %>%
  pull(id.layer_local_c)

lucas.2009 <- lucas.2009 %>%
  filter(!(id.layer_local_c %in% dupli.ids.2009)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "LUCAS 2009/2012; SoilAttr_*.gpkg") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(lucas.2009))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
lucas.2009.trans <- transform_values(df = lucas.2009,
                                     out.name = names(lucas.2009),
                                     in.name = names(lucas.2009),
                                     fun.lst = functions.list)

# Data from 2015 survey
lucas.2015 = fread(path(dir, "/LUCAS_Topsoil_2015_20200323.csv")) %>%
  mutate(Point_ID = paste0("2015.", Point_ID))

# Harmonization of names and units
# 2015 analyte selection and names
analytes.2015.old.names <- transvalues %>%
  filter(table == "LUCAS_Topsoil_2015.csv") %>%
  filter(original_name != "POINT_ID") %>%
  pull(original_name)

analytes.2015.new.names <- transvalues %>%
  filter(table == "LUCAS_Topsoil_2015.csv") %>%
  filter(original_name != "POINT_ID") %>%
  pull(ossl_name)

# Selecting and renaming
lucas.2015 <- lucas.2015 %>%
  rename(id.layer_local_c = Point_ID) %>%
  select(id.layer_local_c, all_of(analytes.2015.old.names)) %>%
  rename_with(~analytes.2015.new.names, all_of(analytes.2015.old.names))

# Removing duplicates
# lucas.2015 %>%
#   group_by(id.layer_local_c) %>%
#   summarise(repeats = n()) %>%
#   group_by(repeats) %>%
#   summarise(count = n())

dupli.ids.2015 <- lucas.2015 %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  filter(repeats > 1) %>%
  pull(id.layer_local_c)

lucas.2015 <- lucas.2015 %>%
  filter(!(id.layer_local_c %in% dupli.ids.2015)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "LUCAS_Topsoil_2015.csv") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(lucas.2015))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
lucas.2015.trans <- transform_values(df = lucas.2015,
                                     out.name = names(lucas.2015),
                                     in.name = names(lucas.2015),
                                     fun.lst = functions.list)

# Final soillab data
lucas.soildata <- bind_rows(lucas.2009.trans, lucas.2015.trans) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
lucas.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.exp.file = path(dir, "ossl_soillab_v1.3")
readr::write_csv(lucas.soildata, str_c(soillab.exp.file, ".csv.gz"))
nanoparquet::write_parquet(lucas.soildata, str_c(soillab.exp.file, ".parquet"))
```

Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
lucas.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### VisNIR spectra

VNIR spectra is stored in absorbacen units, need to transform to reflectance. Also, splice correction at 

```{r visnir, include=TRUE, echo=TRUE, eval=TRUE}
# Spectra from 2009/2012
load(path(dir, "LUCAS.SOIL_corr.Rdata"))

visnir.scans.2009 <- LUCAS.SOIL$spc

lucas.2009.keys <- LUCAS.SOIL %>%
  select(ID, POINT_ID) %>%
  rename(id.scan_local_c = ID, id.layer_local_c = POINT_ID)

visnir.scans.2009 <- visnir.scans.2009 %>%
  rownames_to_column(var = "id.scan_local_c") %>%
  select(id.scan_local_c, any_of(as.character(seq(400, 2500, by = 0.5))))

# Not interpolation to avoid smoothing the splice
old.wavelength <- names(visnir.scans.2009)[-1]
new.wavelength <- seq(400, 2500, by = 2)

visnir.scans.2009 <- visnir.scans.2009 %>%
  select(id.scan_local_c, any_of(as.character(new.wavelength)))
# as.matrix() %>%
# prospectr::resample(X = ., wav = old.wavelength, 
#                     new.wav = new.wavelength, interpol = "spline") %>%
# as_tibble() %>%
# bind_cols({visnir.scans.2009 %>%
#     select(id.scan_local_c)}, .)

# # Metadata 2009/2012
# visnir.metadata.2009 <- visnir.scans.2009 %>%
#   select(id.scan_local_c) %>%
#   mutate(scan.visnir.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2009-05-01"),
#          scan.visnir.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2012-08-01")) %>%
#   mutate(scan.visnir.license.title_ascii_txt = "JRC License Agreement",
#          scan.visnir.license.address_idn_url = "https://esdac.jrc.ec.europa.eu/content/lucas-2009-topsoil-data",
#          scan.visnir.doi_idf_url = "https://data.europa.eu/doi/10.2788/97922")

# lucas.visnir.2009 <- left_join(visnir.metadata.2009, visnir.scans.2009, by = "id.scan_local_c") %>%
#   left_join(lucas.2009.keys, ., by = "id.scan_local_c")

lucas.visnir.2009 <- left_join(visnir.scans.2009, lucas.2009.keys, by = "id.scan_local_c")

lucas.visnir.2009 <- lucas.visnir.2009 %>%
  mutate(id.layer_local_c = paste0("2009.", id.layer_local_c)) %>%
  relocate(id.layer_local_c, .before = 1)

# Spectra from 2015
visnir.scans.2015 <- readRDS(path(dir, "LUCAS_spectra_2015.rds"))

visnir.scans.2015 <- visnir.scans.2015 %>%
  select(PointID, any_of(as.character(seq(400, 2500, by = 0.5)))) %>%
  rename(id.layer_local_c = PointID)

old.wavelength <- names(visnir.scans.2015)[-1]
new.wavelength <- seq(400, 2500, by = 2)

# No interpolation needed
visnir.scans.2015 <- visnir.scans.2015 %>%
  select(id.layer_local_c, any_of(as.character(new.wavelength)))
  # select(-id.layer_local_c) %>%
  # as.matrix() %>%
  # prospectr::resample(X = ., wav = old.wavelength,
  #                     new.wav = new.wavelength, interpol = "spline") %>%
  # as_tibble() %>%
  # bind_cols({visnir.scans.2015 %>%
  #     select(id.layer_local_c)}, .)

# visnir.scans.2015 <- visnir.scans.2015 %>%
#   group_by(id.layer_local_c) %>%
#   summarise_all(mean)

visnir.scans.2015 <- visnir.scans.2015 %>%
  mutate(across(-id.layer_local_c, ~as.numeric(.x)))

visnir.scans.2015 <- visnir.scans.2015[, lapply(.SD, mean), by = id.layer_local_c]

# visnir.metadata.2015 <- visnir.scans.2015 %>%
#   select(id.layer_local_c) %>%
#   mutate(scan.visnir.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2015-03-01"),
#          scan.visnir.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2015-12-01")) %>%
#   mutate(scan.visnir.license.title_ascii_txt = "JRC License Agreement",
#          scan.visnir.license.address_idn_url = "https://esdac.jrc.ec.europa.eu/content/lucas-2015-topsoil-data",
#          scan.visnir.doi_idf_url = "https://data.europa.eu/doi/10.2760/616084")
# 
# lucas.visnir.2015 <- left_join(visnir.metadata.2015, visnir.scans.2015, by = "id.layer_local_c")

lucas.visnir.2015 <- visnir.scans.2015 %>%
  mutate(id.scan_local_c = as.character(id.layer_local_c), .after = id.layer_local_c) %>%
  mutate(id.layer_local_c = paste0("2015.", id.layer_local_c))

# Binding both spectra
# lucas.visnir <- bind_rows(lucas.visnir.2009, lucas.visnir.2015) %>%
#   mutate(id.scan_local_c = id.layer_local_c, .after = id.layer_local_c) %>%
#   mutate(scan.visnir.model.name_utf8_txt = "Metrohm NIRS XDS RapidContent Analyzer",
#          scan.visnir.model.code_any_txt = "Metrohm_NIRS_XDS_RapidContent_Analyzer",
#          scan.visnir.method.optics_any_txt = "",
#          scan.visnir.method.preparation_any_txt = "Sieved <2 mm",
#          scan.visnir.contact.name_utf8_txt = "ESDAC - European Commission",
#          scan.visnir.contact.email_ietf_txt = "ec-esdac@jrc.ec.europa.eu",
#          .after = scan.visnir.doi_idf_url)

# Padding and reflectance transform
lucas.visnir <- bind_rows(lucas.visnir.2009, lucas.visnir.2015)

lucas.visnir <- lucas.visnir %>%
  mutate(`2500` = `2498`)

lucas.visnir <- lucas.visnir %>%
  mutate(across(all_of(as.character(new.wavelength)), ~1/10^(.x)))

# Splice correction
lucas.visnir <- lucas.visnir %>%
  select(-starts_with("id")) %>%
  as.matrix() %>%
  spliceCorrection(wav = new.wavelength,
                   splice = c(1098), interpol.bands = 10) %>%
  as_tibble() %>%
  bind_cols({lucas.visnir %>%
      select(starts_with("id"))}, .)

# Spectral consistency analysis

# Gaps
scans.na.gaps <- lucas.visnir %>%
  dplyr::select(all_of(as.character(new.wavelength))) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({lucas.visnir %>% select(id.scan_local_c)}, .)

# Extreme negative
scans.extreme.neg <- lucas.visnir %>%
  select(all_of(as.character(new.wavelength))) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({lucas.visnir %>% select(id.scan_local_c)}, .)

# Extreme positive
scans.extreme.pos <- lucas.visnir %>%
  select(all_of(as.character(new.wavelength))) %>%
  apply(., 1, function(x) {round(100*(sum(x > 1, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherRef1 = .) %>%
  bind_cols({lucas.visnir %>% select(id.scan_local_c)}, .)

# Consistency summary
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.scan_local_c") %>%
  left_join(scans.extreme.pos, by = "id.scan_local_c")

scans.summary %>%
  select(-id.scan_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming and exporting
old.wavelength <- as.character(seq(400, 2500, by = 2))
new.wavelength <- paste0("scan_visnir.", old.wavelength, "_ref")

lucas.visnir <- lucas.visnir %>%
  rename_with(~new.wavelength, all_of(as.character(old.wavelength)))

lucas.visnir.export <- lucas.visnir %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
visnir.exp.file = path(dir, "ossl_visnir_v1.3")
readr::write_csv(lucas.visnir.export, str_c(visnir.exp.file, ".csv.gz"))
nanoparquet::write_parquet(lucas.visnir.export, str_c(visnir.exp.file, ".parquet"))
```

### Quality control for Vis-NIR

The final table must be joined as follows:

- VisNIR is used as first reference for pairing with soil data.
- Site and soil lab data are left joined to VisNIR. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_visnir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
lucas.availability <- lucas.visnir.export %>%
  select(id.layer_local_c, scan_visnir.600_ref) %>%
  left_join(lucas.soildata, by = "id.layer_local_c")

# Availability of information from lucas
lucas.availability %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())
```

Soil analytical data summary for Vis-NIR. Note: many scans could not be linked with the wetchem.
```{r summary_visnir, include=TRUE, echo=TRUE, eval=TRUE}
lucas.availability %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

Vis-NIR spectral visualization (100 random spectra):
```{r visnir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
lucas.visnir.export %>%
  sample_n(100) %>%
  pivot_longer(starts_with("scan_visnir."),
               names_to = "wavelength", values_to = "intensity") %>%
  mutate(wavelength = gsub("scan_visnir.|_ref", "", wavelength)) %>%
  mutate(wavelength = as.numeric(wavelength),
         intensity = as.numeric(intensity)) %>%
  ggplot(aes(x = wavelength, y = intensity, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  # scale_x_continuous(breaks = seq(350, 2500, by = 250))+
  labs(title = "VisNIR Spectra (100 random scans)",
       x = "Wavelength (nm)",
       y = "Reflectance") +
  theme_light()
```

### Mid-infrared spectra (MIR)

MIR is stored in reflectance units, so we transform to absorbance.
```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
# Floating wavenumbers
mir.scans <- fread(path(dir, "LUCAS_Woodwell.csv"), header = TRUE)

# Need to resample spectra
old.wavenumber <- na.omit(as.numeric(names(mir.scans)))
new.wavenumbers <- rev(seq(628, 4000, by = 2))

lucas.mir <- mir.scans %>%
  select(POINT_ID, all_of(as.character(old.wavenumber))) %>%
  rename(id.layer_local_c = POINT_ID) %>%
  mutate(id.layer_local_c = paste0("2009.", id.layer_local_c))

lucas.mir <- lucas.mir %>%
  select(-id.layer_local_c) %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavenumber,
                      new.wav = new.wavenumbers, interpol = "spline") %>%
  as_tibble() %>%
  bind_cols({lucas.mir %>%
      select(id.layer_local_c)}, .) %>%
  select(id.layer_local_c, as.character(rev(new.wavenumbers))) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# Checking duplicates
dupli.ids <- lucas.mir %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  filter(repeats > 1) %>%
  pull(id.layer_local_c)

lucas.mir <- lucas.mir %>%
  filter(!(id.layer_local_c %in% dupli.ids)) %>%
  as_tibble()

# Spectral consistency analysis

# Gaps
scans.na.gaps <- lucas.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({lucas.mir %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
scans.extreme.neg <- lucas.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < -1, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({lucas.mir %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
scans.extreme.pos <- lucas.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 3.5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({lucas.mir %>% select(id.layer_local_c)}, .)

# Consistency summary - problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# # These few scans with extreme values are filled with interpolation
# target.wavenumbers <- seq(600, 4000, by = 2)
# 
# lucas.mir.clean <- lucas.mir %>%
#   pivot_longer(-id.layer_local_c, names_to = "wavenumber", values_to = "absorbance") %>%
#   mutate(absorbance = ifelse(absorbance > 3.5, NA, absorbance)) %>%
#   group_by(id.layer_local_c) %>%
#   mutate(absorbance = approx(wavenumber, absorbance, target.wavenumbers)$y) %>%
#   pivot_wider(names_from = "wavenumber", values_from = "absorbance")

# Renaming
new.wavenumbers.names <- paste0("scan_mir.", new.wavenumbers, "_abs")

lucas.mir <- lucas.mir %>%
  rename_with(~new.wavenumbers.names, as.character(new.wavenumbers))

# # Preparing metadata
# lucas.mir.metadata <- mir.scans %>%
#   select(POINT_ID, WHRC_ID, run_date) %>%
#   rename(id.layer_local_c = POINT_ID, id.scan_local_c = WHRC_ID,
#          scan.mir.date.begin_iso.8601_yyyy.mm.dd = run_date) %>%
#   mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
#   mutate(id.layer_local_c = paste0("2009.", id.layer_local_c),
#          scan.mir.date.begin_iso.8601_yyyy.mm.dd = dmy(scan.mir.date.begin_iso.8601_yyyy.mm.dd)) %>%
#   filter(!(id.layer_local_c %in% dupli.ids)) %>%
#   mutate(dataset.code_ascii_txt = "LUCAS.WOODWELL.SSL",
#          scan.mir.date.end_iso.8601_yyyy.mm.dd = scan.mir.date.begin_iso.8601_yyyy.mm.dd,
#          scan.mir.model.name_utf8_txt = "Bruker Vertex 70 with with PikeAutoDiff accessory",
#          scan.mir.model.code_any_txt = "Bruker_Vertex_70.PikeAutoDiff",
#          scan.mir.method.optics_any_txt = "KBr beamsplitter; Gold mirror; Mirror background",
#          scan.mir.method.preparation_any_txt = "",
#          scan.mir.license.title_ascii_txt = "CC-BY",
#          scan.mir.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/",
#          scan.mir.doi_idf_url = "https://doi.org/10.3390/s20236729",
#          scan.mir.contact.name_utf8_txt = "Jonathan Sanderman",
#          scan.mir.contact.email_ietf_txt = "jsanderman@woodwellclimate.org")

# Final preparation
# lucas.mir.export <- lucas.mir.metadata %>%
#   left_join(lucas.mir, by = "id.layer_local_c") %>%
#   mutate_at(vars(starts_with("id.")), as.character)

lucas.mir.export <- lucas.mir %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
mir.exp.file = path(dir, "ossl_mir_v1.3")
readr::write_csv(lucas.mir.export, str_c(mir.exp.file, ".csv.gz"))
nanoparquet::write_parquet(lucas.mir.export, str_c(mir.exp.file, ".parquet"))
```

### Quality control for MIR

The final table must be joined as follows:

- MIR is used as first reference for pairing with soil data.
- Soil lab data are left joined to MIR. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_mir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
lucas.availability2 <- lucas.mir.export %>%
  select(id.layer_local_c, scan_mir.1000_abs) %>%
  left_join(lucas.soildata, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from lucas
lucas.availability2 %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Duplicates are dropped
lucas.availability2 %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

Soil analytical data summary for MIR. Note: many scans could not be linked with the wetchem.
```{r summary_mir, include=TRUE, echo=TRUE, eval=TRUE}
lucas.availability2 %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
lucas.mir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "reflectance") %>%
  dplyr::mutate(wavenumber = gsub("scan_mir.|_abs", "", wavenumber)) %>%
  dplyr::mutate(wavenumber = as.numeric(wavenumber)) %>%
  ggplot(aes(x = wavenumber, y = reflectance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = c(600, 1200, 1800, 2400, 3000, 3600, 4000),
                     transform = "reverse") +
  labs(title = "MIR Spectra (100 random scans)",
       x = bquote("Wavenumber"~(cm^-1)),
       y = "Absorbance")+
  theme_light()
```

```{r end}
toc()
rm(list = ls())
gc()
```

## References
