---
title: "Geocradle dataset preparation for the OSSL"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
date: last-modified
date-format: "--- DD MMMM, YYYY."
format: 
  gfm:
    toc: true
    toc-depth: 2
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Geocradle dataset into the Open Soil Spectral Library.

Project: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The Geocradle original data

Site data, Soil lab data, and Visible Near-Infrared (Vis-NIR) data from the Geocradle regional soil spectral library. Further information of the dataset can be
found in detail at @tziolas_integrated_2020 and @tziolas_memory-based_2019.

Original files:\
- `SSL_GEOCRADLE_1.csv`: csv file with site information, soil information, and Vis-NIR spectral data.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "prospectr", "units", "readxl", "stringr",
              "olctools", "openssl", "tmap", "sf", "skimr", "lubridate",
              "googledrive", "googlesheets4", "data.table", "doMC", "tictoc",
              "qs", "dplyr", "tidyr", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r}
dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/Geocradle"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
geocradle.metadata <- fread(path(dir, "SSL_GEOCRADLE_1.csv"), header = T)

geocradle.sitedata <- geocradle.metadata %>%
  select(ID, origin, Latitude, Longitude, Sampling_date, Elevation, Depth, Soil_type_WRB, Climate_Koeppen) %>%
  rename(id.layer_local_c = ID,
         longitude.point_wgs84_dd = Longitude,
         latitude.point_wgs84_dd = Latitude) %>%
  mutate(id.project_ascii_txt = "Geocradle Regional Soil Spectral Library",
         dataset.code_ascii_txt = "GEOCRADLE.SSL",
         observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
         observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
         dataset.title_utf8_txt = "GEOCRADLE: Geocradle Regional Soil Spectral Library for Balkans, Middle East, and North Africa",
         dataset.owner_utf8_txt = "PILOT 2: Improved Food Security – Water Extremes Management (IFS)",
         dataset.doi_idf_url = "http://datahub.geocradle.eu/dataset/regional-soil-spectral-library",
         dataset.license.title_ascii_txt = "Open Data Commons Open Database License (ODbL)",
         dataset.license.address_idn_url = "https://opendefinition.org/licenses/odc-odbl/",
         dataset.contact.name_utf8_txt = "Nikos Tsakiridis",
         dataset.contact_ietf_email = "tsakirin@auth.gr") %>%
  mutate(id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         id.location_olc_txt = olctools::encode_olc(latitude.point_wgs84_dd, longitude.point_wgs84_dd, 10),
         .after = id.project_ascii_txt) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.qs = path(dir, "ossl_soilsite_v2.0.qs")
qs::qsave(geocradle.sitedata, site.qs, preset = "high")
```

Plotting sites map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

points <- geocradle.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")

tm_shape(World) +
  tm_polygons('#f0f0f0f0', border.alpha = 0.2) +
  tm_shape(points) +
  tm_dots(col = "origin", 
          palette = "Set1") +
  tm_layout(legend.outside = TRUE)

```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwritting, then rename online to download locally.**

```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- geocradle.metadata %>%
  select(ID, Sand_Fraction, Clay_Fraction, Silt_Fraction, USDA_texture, OC, OM, Soil_type_extended_WRB, Soil_type_USDA,
         Soil_extended_WRB, Soil_type_WRB_description, CaCO3, CEC, LOI, pH_H2O, pH_KCl, pH_CaCl2, EC_muS, NO3) %>%
  rename(id.layer_local_c = ID) %>%
  names() %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'SSL_GEOCRADLE_1.csv', .before = 1) %>%
  dplyr::mutate(import = '', original_unit = '', comment = '', ossl_abbrev = '', ossl_method = '', ossl_unit = '',
                ossl_convert = '', ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the output file is saved to this folder for archiving purposes.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "Geocradle") %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
geocradle.reference <- geocradle.metadata

# Harmonization of names and units

valid_mappings <- transvalues %>%
  filter(table == "SSL_GEOCRADLE_1.csv") %>%
  filter(!is.na(ossl_name) & ossl_name != "") 

analytes.old.names <- valid_mappings %>% pull(original_name)
analytes.new.names <- valid_mappings %>% pull(ossl_name)

analytes.old.names.clean <- analytes.old.names[analytes.old.names != "ID"]
analytes.new.names.clean <- analytes.new.names[analytes.old.names != "ID"]

geocradle.soildata <- geocradle.reference %>%
  rename(id.layer_local_c = ID) %>%
  select(id.layer_local_c, all_of(analytes.old.names.clean)) %>%
  rename_with(~analytes.new.names.clean, all_of(analytes.old.names.clean)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "SSL_GEOCRADLE_1.csv") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(geocradle.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
geocradle.soildata.trans <- transform_values(df = geocradle.soildata,
                                       out.name = names(geocradle.soildata),
                                       in.name = names(geocradle.soildata),
                                       fun.lst = functions.list)

# Final soillab data
geocradle.soildata <- geocradle.soildata.trans %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  mutate(id.layer_local_c = make.unique(id.layer_local_c))

# Checking total number of observations
geocradle.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())


# Saving version to dataset root dir
soillab.qs = path(dir, "ossl_soillab_v2.0.qs")
qs::qsave(geocradle.soildata, soillab.qs, preset = "high")
```

Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
geocradle.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### Vis-NIR spectra

```{r vis-nir, include=TRUE, echo=TRUE, eval=TRUE}
# Renaming and ID Preparation
geocradle.visnir.proc <- geocradle.metadata %>%
  as_tibble() %>%
  rename(id.layer_local_c = ID) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# Extracting wavelengths from column names
spec.cols <- grep("^X[0-9].", names(geocradle.visnir.proc), value = TRUE)
old.wavelengths <- as.numeric(gsub("X", "", spec.cols))
new.wavelengths <- seq(350, 2500, by = 2)

# Resampling spectra
geocradle.visnir.proc <- geocradle.visnir.proc %>%
  select(id.layer_local_c, all_of(spec.cols)) %>%
  { . } -> temp_data 

geocradle.visnir.resampled <- temp_data %>%
  mutate(id.layer_local_c = make.unique(as.character(id.layer_local_c))) %>%
  {
    spec_matrix <- . %>% 
      select(all_of(spec.cols)) %>% 
      as.matrix()
    ids <- . %>% select(id.layer_local_c)
    resampled_mat <- prospectr::resample(
      X = spec_matrix(.), 
      wav = old.wavelengths, 
      new.wav = new.wavelengths, 
      interpol = "spline"
    )
    bind_cols(ids(.), as_tibble(resampled_mat))
  } %>%
  select(id.layer_local_c, all_of(as.character(new.wavelengths)))

# Gaps Analysis
scans.na.gaps <- geocradle.visnir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols(geocradle.visnir.resampled %>% select(id.layer_local_c), .)

# Extreme negative checks
scans.extreme.neg <- geocradle.visnir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols(geocradle.visnir.resampled %>% select(id.layer_local_c), .)

# Extreme positive checks
scans.extreme.pos <- geocradle.visnir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 1.5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols(geocradle.visnir.resampled %>% select(id.layer_local_c), .)

# Consistency summary
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

# Display problematic scans count
scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Final column renaming for OSSL standard
final.visnir.names <- paste0("scan_visnir.", new.wavelengths, "_ref")
geocradle.visnir.proc <- geocradle.visnir.resampled %>%
  rename_with(~final.visnir.names, as.character(new.wavelengths))

# Preparing metadata
geocradle.visnir.metadata <- geocradle.visnir.proc %>%
  select(id.layer_local_c) %>%
  mutate(id.scan_local_c = id.layer_local_c,
         scan.visnir.date.begin_iso.8601_yyyy = ymd("2017-01-01"), 
         scan.visnir.date.end_iso.8601_yyyy = ymd("2017-12-31"), 
         scan.visnir.model.name_utf8_txt = "ASD Fieldspec Pro (PANanalytical B·V, Boulder, CO, USA formerly Analytical Spectral Devices, at IS, CY and TR) and a PSR + spectrometer (Spectral Evolution Inc., Lawrence, Massachusetts, USA, at GR, AL, FY, BG, EG, and RS)", 
         scan.visnir.license.title_ascii_txt = "Open Data Commons Open Database License (ODbL)",
         scan.visnir.method.optics_any_txt = "",
         scan.visnir.method.preparation_any_txt = "Standard preparation",
         scan.visnir.license.address_idn_url = "https://opendefinition.org/licenses/odc-odbl/",
         scan.visnir.doi_idf_url = "http://datahub.geocradle.eu/dataset/regional-soil-spectral-library",
         scan.visnir.contact.name_utf8_txt = "Nikos Tsakiridis",
         scan.visnir.contact.email_ietf_txt = "tsakirin@auth.gr")

# Final preparation
 geocradle.visnir.export <- geocradle.visnir.metadata %>%
  left_join(geocradle.visnir.proc, by = "id.layer_local_c") %>%
  mutate(across(starts_with("id."), as.character))

# Saving
soilvisnir.qs = path(dir, "ossl_visnir_v2.0.qs")
qs::qsave(geocradle.visnir.export, soilvisnir.qs, preset = "high")

```


### Quality control for Vis-NIR

The final table must be joined as follows:

- Vis-NIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_visnir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins

metadata_clean <- geocradle.metadata %>%
  rename(id.layer_local_c = ID) %>%
  mutate(id.layer_local_c = make.unique(as.character(id.layer_local_c))) %>%
  select(id.layer_local_c, Latitude, Sand_Fraction, pH_H2O)

geocradle.availability <- geocradle.visnir.export %>%
  select(id.layer_local_c, scan_visnir.600_ref) %>%
  left_join(metadata_clean, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))


# Availability of information summary
# This tells us how many samples have spectra vs. lab/site data
geocradle.availability %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Checking for duplicate IDs
geocradle.availability %>%
  mutate(across(everything(), as.character)) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n(), .groups = 'drop') %>%
  group_by(column, repeats) %>%
  summarise(count = n(), .groups = 'drop')

```

Soil analytical data summary for Vis-NIR. Note: many scans could not be linked with the wetchem.
```{r summary_visnir, include=TRUE, echo=TRUE, eval=TRUE}

all(geocradle.soildata$id.layer_local_c %in% geocradle.visnir.export$id.layer_local_c)
geocradle.soildata %>%
  filter(id.layer_local_c %in% geocradle.visnir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```
Vis-NIR spectral visualization (100 random spectra):
```{r visnir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
geocradle.visnir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_visnir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavelength", values_to = "reflectance") %>%
  dplyr::mutate(wavelength = gsub("scan_visnir.|_ref", "", wavelength)) %>%
  dplyr::mutate(wavelength = as.numeric(wavelength)) %>%
  ggplot(aes(x = wavelength, y = reflectance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(350, 2500, by = 250))+
  labs(title = "Vis-NIR Spectra (100 random scans)",
       x = "Wavelength (nm)",
       y = "Reflectance")+
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```

## References
