---
title: "Southwestern India dataset preparation for the OSSL"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org

date: last-modified
date-format: "--- DD MMMM, YYYY."
format: 
  gfm:
    toc: true
    toc-depth: 2
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Southwestern India dataset into the Open Soil Spectral Library.

Project: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The Southwestern India original data

Site data, Soil lab data, and Mid-Infrared Spectra (MIR) from small-scale tropical, sub-humid and semi-arid watersheds under shrubland and dry deciduous forest in Southwestern India. Further information of the dataset can be found in detail at @belle_soil_2022.

Original files:\
- `20211111_MS_SOC_Stock_India_data_raw.csv`: csv file with site information and soil information. 
- `20211111_DRFT_data`: a folder has MIR spectral data through Diffuse Reflectance Infrared Fourier-Transformed (DRIFT) spectroscopy.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "prospectr", "units", "readxl", "stringr",
              "olctools", "openssl", "tmap", "sf", "skimr", "lubridate",
              "googledrive", "googlesheets4", "data.table", "doMC", "tictoc",
              "qs", "dplyr", "tidyr", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r}
dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/South India"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
india.rawdata <- fread(path(dir, "20211111_MS_SOC_Stock_India_data_raw.csv"), header = T)
india.compdata <- fread(path(dir, "20211111_MS_SOC_Stock_India_data_composites.csv"), header = T)

india.metadata <- india.rawdata %>%
  left_join(india.compdata, by = c("Site", "Depth")) %>%
  mutate(sample_id = paste(Site, Pit, Depth, sep = "_")) 

india.sitedata <- india.metadata %>%
  select(sample_id, Site, Pit, Depth, ) %>%
  separate(Depth, 
           into = c("layer.upper.depth_usda_cm", "layer.lower.depth_usda_cm"), 
           sep = "-", 
           convert = TRUE, 
           remove = FALSE) %>%
  rename(id.layer_local_c = sample_id) %>%
  mutate(across(id.layer_local_c, as.character)) %>%
  mutate(id.project_ascii_txt = "South India Soil Spectral Library",
         dataset.code_ascii_txt = "SouthIndia.MIR",
         observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
         observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
         surveyor.title_utf8_txt = "Severin-Luca Bell`e, Samuel Abiven",
         surveyor.contact_ietf_email = "abiven@biotite.ens.fr",
         dataset.title_utf8_txt = "South India Soil Spectral Library",
         dataset.owner_utf8_txt = "Samuel Abiven",
         dataset.address_idn_url = "https://zenodo.org/records/5675793",
         dataset.doi_idf_url = "https://zenodo.org/records/5675793",
         dataset.license.title_ascii_txt = "CC-BY 4.0",
         dataset.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/legalcode",
         dataset.contact.name_utf8_txt = "Samuel Abiven",
         dataset.contact_ietf_email = "abiven@biotite.ens.fr") %>%
  mutate(id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt)),
         .after = id.project_ascii_txt) %>%
  mutate_at(vars(starts_with("id.")), as.character)

write_csv(india.metadata, "india_metadata.csv")
# Saving version to dataset root dir
site.qs = path(dir, "ossl_soilsite_v2.0.qs")
qs::qsave(india.sitedata, site.qs, preset = "high")
```


### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.


```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- india.metadata %>%
  select(sample_id, BD, TC, TOC, `SOC stock`, TN, CN, pH, EC, CEC, Clay, Silt, Sand, 
         Fetot, Altot, Sitot, Ktot, Natot, Catot, Ptot, Mgtot) %>%
  names(.) %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'india_metadata.csv', .before = 1) %>%
  dplyr::mutate(import = '', original_unit = '',  original_method = '', comment = '', 
                ossl_abbrev = '', ossl_method = '', ossl_unit = '', ossl_convert = '', 
                ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the output file is saved to this folder for archiving purposes.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "South-India") %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
india.reference <- india.metadata

# Harmonization of names and units
analytes.old.names <- transvalues %>%
  filter(table == "india_metadata.csv") %>%
  pull(original_name)

analytes.new.names <- transvalues %>%
  filter(table == "india_metadata.csv") %>%
  pull(ossl_name)

# Selecting and renaming
analytes.old.names.clean <- analytes.old.names[analytes.old.names != "sample_id"]
analytes.new.names.clean <- analytes.new.names[analytes.old.names != "sample_id"]

india.soildata <- india.reference %>%
  rename(id.layer_local_c = sample_id) %>%
  select(id.layer_local_c, all_of(analytes.old.names.clean)) %>%
  rename_with(~analytes.new.names.clean, all_of(analytes.old.names.clean)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "india_metadata.csv") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(india.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
india.soildata.trans <- transform_values(df = india.soildata,
                                       out.name = names(india.soildata),
                                       in.name = names(india.soildata),
                                       fun.lst = functions.list)

# Final soillab data
india.soildata <- india.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
india.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.qs = path(dir, "ossl_soillab_v2.0.qs")
qs::qsave(india.soildata, soillab.qs, preset = "high")
```

Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
india.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```


### Mid-infrared spectra (MIR)
```{r mir_load, include=TRUE, echo=TRUE, eval=TRUE}
## Firstly, read MIR data
folder_path <- "20211111_DRIFT_data"
file_list <- dir_ls(folder_path, glob = "*.dpt")

# Function to read a single .dpt file
read_dpt <- function(file_path) {
  data <- read_csv(file_path, col_names = c("wavenumber", "absorbance"), 
                   col_types = cols(wavenumber = col_double(), absorbance = col_double()),
                   show_col_types = FALSE)
  ID_val <- path_file(file_path) %>% str_replace(".dpt", "")
  data %>%
    mutate(ID = ID_val)
}

mir_long <- file_list %>%
  map_df(~read_dpt(.x))

india.mir <- mir_long %>%
  mutate(wavenumber = paste0("scan_MIR_", round(wavenumber))) %>%
  pivot_wider(names_from = wavenumber, values_from = absorbance)
  
## Assign sample id to MIR data
mir_id_map <- read_delim(file.path(folder_path, "20211111_DRIFT_ID_all.csv"), delim = ";", show_col_types = FALSE)

mir_id_map <- mir_id_map %>%
  mutate(sample_id = paste(Site, Pit, Increment, sep = "_"),
         ID = as.character(ID))

india.mir <- india.mir %>%
  mutate(ID = as.character(ID)) %>%
  inner_join(mir_id_map %>% select(ID, sample_id), by = "ID") %>%
  select (sample_id, everything())

## Combine MIR with metadata
india.final <- inner_join(india.metadata, india.mir, by = "sample_id") %>%
  relocate(sample_id, .before =1) %>%
  select( - ID)

write_csv(india.final, "india_fulldata.csv")
```


```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
# Renaming
india.mir.proc <- india.final %>%
  rename(id.layer_local_c = sample_id) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

mir_cols <- names(india.mir.proc)[grep("scan_MIR_", names(india.mir.proc))]

# Resample spectra
old.wavenumber.num <- as.numeric(str_replace(mir_cols, "scan_MIR_", ""))
new.wavenumber <- seq(600, 4000, by = 2)

india.mir.matrix <- india.mir.proc %>%
  select(all_of(mir_cols)) %>%
  as.matrix()

india.mir.resampled <- prospectr::resample(
  X = india.mir.matrix, 
  wav = old.wavenumber.num, 
  new.wav = new.wavenumber, 
  interpol = "spline"
) %>%
  as_tibble()

india.mir.proc <- bind_cols(
  india.mir.proc %>% select(id.layer_local_c), 
  india.mir.resampled
)

# Gaps

mir.na.gaps <- india.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({india.mir.proc %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
mir.extreme.neg <- india.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({india.mir.proc %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
mir.extreme.pos <- india.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({india.mir.proc %>% select(id.layer_local_c)}, .)


# Consistency summary - problematic scans
mir.summary <- mir.na.gaps %>%
  left_join(mir.extreme.neg, by = "id.layer_local_c") %>%
  left_join(mir.extreme.pos, by = "id.layer_local_c")

mir.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
final.mir.names <- paste0("scan_mir.", new.wavenumber, "_abs")
names(india.mir.resampled) <- final.mir.names

india.mir.proc <- bind_cols(
  india.mir.proc %>% select(id.layer_local_c),
  india.mir.resampled
)

# Preparing metadata
india.mir.metadata <- india.mir.proc %>%
  select(id.layer_local_c) %>%
  mutate(id.scan_local_c = id.layer_local_c,
         scan.mir.model.name_utf8_txt = "TENSOR 27 spectrophotometer, Bruker, Switzerland", 
         scan.mir.license.title_ascii_txt = "CC-BY 4.0",
         scan.mir.method.optics_any_txt = "",
         scan.mir.method.preparation_any_txt = "All pit core samples were dried at 40◦C for 48 h and subsequently
sieved to < 2 mm. Then samples were milled using a horizontal mill (2–5 min, 30 turns/
s).",
         scan.mir.doi_idf_url = "https://zenodo.org/records/5675793",
         scan.mir.contact.name_utf8_txt = "Samuel Abiven",
         scan.mir.contact.email_ietf_txt = "abiven@biotite.ens.fr")

# Final preparation
india.mir.export <- india.mir.metadata %>%
  left_join(india.mir.proc, by = "id.layer_local_c") %>%
  mutate(across(starts_with("id."), as.character))

# Saving version to dataset root dir
soilmir.qs = path(dir, "ossl_mir_v2.0.qs")
qs::qsave(india.mir.export, soilmir.qs, preset = "high")
```

### Quality control 

The final table must be joined as follows:

- MIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_mir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
india.availability <- india.mir.export %>%
  select(id.layer_local_c, scan_mir.600_abs) %>%
  left_join({india.sitedata %>%
      select(id.layer_local_c)}, by = "id.layer_local_c") %>%
  left_join({india.soildata %>%
      select(id.layer_local_c, c.tot_iso.10694_w.pct)}, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from besb
india.info.summary <- india.availability %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())
print("Availability Summary:")
print(india.info.summary)

# Repeats check - Duplicates are dropped
india.availability %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

Soil analytical data summary for MIR. Note: many scans could not be linked with the wetchem.
```{r summary_mir, include=TRUE, echo=TRUE, eval=TRUE}
india.soildata %>%
  filter(id.layer_local_c %in% india.mir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
india.mir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "absorbance") %>%
  dplyr::mutate(wavenumber = as.numeric(gsub("scan_mir.|_abs", "", wavenumber))) %>%
  ggplot(aes(x = wavenumber, y = absorbance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(600, 4000, by = 400),
                     transform = "reverse") +
  labs(title = "MIR Spectra (100 random scans)",
       x = bquote("Wavenumber"~(cm^-1)),
       y = "Absorbance")+
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```
## References

