---
title: "Africa Soil Information Service Phase I (AFSIS1) preparation for the OSSL"
author: 
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Ran Zhi
    email: jsafanelli@woodwellclimate.org
  - name: Tomislav Hengl
    email: tom.hengl@opengeohub.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: ../../tex/refs.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for standardizing and importing the ICRAF Soil and Plant Spectroscopy Laboratory: Africa Soil Information Service Phase I 2009-2013 (AFSIS1) into the Open Soil Spectral Library.

Website: [Soil Spectroscopy for Global Good](https://soilspectroscopy.org)  \
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The AFIS1 original data

Site data, soil lab data, and Mid-Infrared Spectra (MIR) from AFSIS1.

Further information can be found in @Towett2015, @Vagen2016, and @Vagen2020.

Input datasets:  
- `AfSIS_reference.csv`: Database with site and soil analytes.\
- `afsis_mir_2013`: folder with MIR soil spectral data.\
- `Calibration_MPA_NIR.csv`: VNIR soil spectral data.\

The dataset has two versions one published via the [World Agroforestry website](https://doi.org/10.34725/DVN/QXCWP1), one via AWS (<https://registry.opendata.aws/afsis/>).

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "data.table", "lubridate", "units", "readxl", "stringr",
              "googledrive", "googlesheets4", "tmap", "sf", "rnaturalearth",
              "skimr", "prospectr", "tictoc", "nanoparquet", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
gs4_auth(path = "~/secrets/soilcarbon-soilspec-845d0d3f2fbe.json")
options(scipen = 999)
```

Directory/folder path with original files (not uploaded to GitHub). 
```{r}
# dir = "./"
dir = "~/mnt-ossl-private/database/datasets/AFSIS1"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
# AfIS1 site data
afsis1.geo <- fread(paste0(dir, "/georeferences_2013.csv"))

afsis1.sitedata <- afsis1.geo %>%
  select(SSN, Latitude, Longitude, Depth,
         Scientist, Cluster, Plot, Site, Country, Region, Cultivated) %>%
  rename(id.layer_local_c = SSN,
         longitude.point_wgs84_dd = Longitude,
         latitude.point_wgs84_dd = Latitude,
         surveyor.title_utf8_txt = Scientist,
         site.cluster_src_id = Cluster,
         site.plot_src_id = Plot,
         site.name_src_txt = Site,
         site.cultivated_src_logical = Cultivated,
         loc.country_src_txt = Country,
         loc.region_src_txt = Region) %>%
  mutate(observation.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2011-01-01"),
         observation.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2013-12-31")) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c),
         layer.upper.depth_usda_cm = ifelse(Depth == "top", 0, 20),
         layer.lower.depth_usda_cm = ifelse(Depth == "top", 20, 50),
         layer.sequence_usda_uint16 = ifelse(Depth == "top", 1, 2)) %>%
  select(-Depth) %>%
  # mutate(id.project_ascii_txt = "Africa Soil Information Service (AfSIS-1)",
  #        layer.texture_usda_txt = "",
  #        pedon.taxa_usda_txt = "",
  #        horizon.designation_usda_txt = "",
  #        longitude.county_wgs84_dd = NA,
  #        latitude.county_wgs84_dd = NA,
  #        location.point.error_any_m = 30,
  #        location.country_iso.3166_txt = "",
  #        observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
  #        observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
  #        surveyor.contact_ietf_email = "afsis.info@africasoils.net",
  #        surveyor.address_utf8_txt = "ICRAF, PO Box 30677, Nairobi, 00100, Kenya",
  #        dataset.title_utf8_txt = "Africa Soil Information Service (AfSIS-1)",
  #        dataset.owner_utf8_txt = "ICRAF, CROPNUTS, RRES",
  #        dataset.code_ascii_txt = "AFSIS1.SSL",
  #        dataset.address_idn_url = "https://www.isric.org/explore/ISRIC-collections",
  #        dataset.doi_idf_url = "https://doi.org/10.1016/j.geodrs.2015.06.002",
  #        dataset.license.title_ascii_txt = "ODC Open Database License",
  #        dataset.license.address_idn_url = "https://opendatacommons.org/licenses/odbl/",
  #        dataset.contact.name_utf8_txt = "Keith Shepherd",
  #        dataset.contact_ietf_email = "afsis.info@africasoils.net") %>%
  mutate(dataset.code_ascii_txt = "AFSIS1.SSL",
         id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .before = 1)

# Removing duplicates
afsis1.sitedata %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  group_by(repeats) %>%
  summarise(count = n())

dupli.ids <- afsis1.sitedata %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  filter(repeats > 1) %>%
  pull(id.layer_local_c)

afsis1.sitedata <- afsis1.sitedata %>%
  filter(!(id.layer_local_c %in% dupli.ids))

# # Removing rows without spatial coordinates
# afsis1.sitedata <- afsis1.sitedata %>%
#   filter(!(is.na(longitude.point_wgs84_dd) & is.na(latitude.point_wgs84_dd))) %>%
#   mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.exp.file = path(dir, "ossl_soilsite_v1.3")
readr::write_csv(afsis1.sitedata, str_c(site.exp.file, ".csv.gz"))
nanoparquet::write_parquet(afsis1.sitedata, str_c(site.exp.file, ".parquet"))
```

Plotting AFSIS1 map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

ocean <- ne_download(scale = 110, type = "ocean", category = "physical", returnclass = "sf")

points <- afsis1.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")
tm_shape(ocean) +
  tm_polygons(fill = "lightblue", col = NA) +
  tm_shape(World) +
  tm_polygons(fill = "#f0f0f0", fill_alpha = 0.5, col_alpha = 0.5) +
  tm_shape(points) +
  tm_dots(fill = "firebrick") +
  tm_crs("ESRI:54030") +
  tm_layout(frame = FALSE)
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template for scripted column standardization. Just run once for getting the original names of soil properties, descriptions, data types, and units. Then upload to Google Sheet for editing and manually defining the rules for integrating with the OSSL. Requires Google authentication. A copy of the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwritting, then rename online to download locally.**

```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables
variables.file = readr::read_csv(paste0(dir, "/AfSIS_reference.csv"))

# They all have the same soil properties
soillab.names <- variables.file %>%
  names(.) %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'AfSIS_reference.csv', .before = 1) %>%
  dplyr::mutate(import = '', ossl_name = '', .after = original_name) %>%
  dplyr::mutate(comment = '')

readr::write_csv(soillab.names, paste0(getwd(), "/afis_soillab_names.csv"))

# Uploading to google sheet

# FACT CIN folder. Get ID for soildata importing table
googledrive::drive_ls(as_id("0AHDIWmLAj40_Uk9PVA"))

OSSL.soildata.importing <- "19LeILz9AEnKVK7GK0ZbK3CCr2RfeP-gSWn5VpY8ETVM"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names
upload <- dplyr::as_tibble(soillab.names)

# Uploading
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the edited standardization template is saved to this dataset folder.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "AFSIS1")

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:
```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F) %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "ossl_")))

knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:
```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
afsis1.reference = fread(paste0(dir, "/AfSIS_reference.csv"))

# Harmonization of names and units
analytes.old.names <- transvalues %>%
  filter(table == "AfSIS_reference.csv") %>%
  pull(original_name)

analytes.new.names <- transvalues %>%
  filter(table == "AfSIS_reference.csv") %>%
  pull(ossl_name)

# Selecting and renaming
afsis1.soildata <- afsis1.reference %>%
  rename(id.layer_local_c = SSN) %>%
  select(id.layer_local_c, all_of(analytes.old.names)) %>%
  rename_with(~analytes.new.names, all_of(analytes.old.names))

# Removing duplicates
# afsis1.soildata %>%
#   group_by(id.layer_local_c) %>%
#   summarise(repeats = n()) %>%
#   group_by(repeats) %>%
#   summarise(count = n())

dupli.ids <- afsis1.soildata %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  filter(repeats > 1) %>%
  pull(id.layer_local_c)

afsis1.soildata <- afsis1.soildata %>%
  filter(!(id.layer_local_c %in% dupli.ids)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "AfSIS_reference.csv") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(afsis1.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
afsis1.soildata.trans <- transform_values(df = afsis1.soildata,
                                          out.name = names(afsis1.soildata),
                                          in.name = names(afsis1.soildata),
                                          fun.lst = functions.list)

# Final soillab data
afsis1.soildata <- afsis1.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
afsis1.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())

# Saving version to dataset root dir
soillab.exp.file = path(dir, "ossl_soillab_v1.3")
readr::write_csv(afsis1.soildata, str_c(soillab.exp.file, ".csv.gz"))
nanoparquet::write_parquet(afsis1.soildata, str_c(soillab.exp.file, ".parquet"))
```

Soil lab data summary.
```{r soil_lab_summary, include=TRUE, echo=TRUE, eval=TRUE}
afsis1.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### Mid-infrared spectra

Spectra already in pseudo absorbance units.

**PLEASE NOTE: Only ~10% of the spectra have paired analytical data.**

```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
# Floating wavenumbers
mir.files <- list.files(paste0(dir, "/afsis_mir_2013"), pattern = "\\.csv$", full.names = T)

mir.scans <- mir.files %>%
  purrr::map_dfr(fread, header = TRUE) %>%
  select(SSN, starts_with("m"))

old.names <- names(mir.scans)
new.names <- gsub("m", "", old.names)

afsis1.mir <- mir.scans %>%
  rename_with(~new.names, all_of(old.names)) %>%
  rename(id.layer_local_c = SSN)

# Removing samples without spectra
scans.na.gaps <- afsis1.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({afsis1.mir %>% select(id.layer_local_c)}, .)

issue.ids <- scans.na.gaps %>%
  filter(proportion_NA > 0) %>%
  pull(id.layer_local_c)

afsis1.mir <- afsis1.mir %>%
  filter(!(id.layer_local_c %in% issue.ids))

# Need to resample spectra
old.wavenumber <- na.omit(as.numeric(names(afsis1.mir)))
new.wavenumbers <- rev(seq(600, 4000, by = 2))

afsis1.mir <- afsis1.mir %>%
  select(-id.layer_local_c) %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavenumber,
                      new.wav = new.wavenumbers, interpol = "spline") %>%
  as_tibble() %>%
  bind_cols({afsis1.mir %>%
      select(id.layer_local_c)}, .) %>%
  select(id.layer_local_c, as.character(rev(new.wavenumbers)))

# Gaps
scans.na.gaps <- afsis1.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({afsis1.mir %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
scans.extreme.neg <- afsis1.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < -1, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({afsis1.mir %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
scans.extreme.pos <- afsis1.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({afsis1.mir %>% select(id.layer_local_c)}, .)

# Consistency summary - problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
old.wavenumbers <- seq(600, 4000, by = 2)
new.wavenumbers <- paste0("scan_mir.", old.wavenumbers, "_abs")

afsis1.mir <- afsis1.mir %>%
  rename_with(~new.wavenumbers, all_of(as.character(old.wavenumbers))) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# # Preparing metadata
# afsis1.mir.metadata <- afsis1.mir %>%
#   select(id.layer_local_c) %>%
#   mutate(id.scan_local_c = id.layer_local_c) %>%
#   mutate(scan.mir.date.begin_iso.8601_yyyy.mm.dd = ymd("2009-01-01"),
#          scan.mir.date.end_iso.8601_yyyy.mm.dd = ymd("2013-12-31"),
#          scan.mir.model.name_utf8_txt = "Bruker Tensor 27 with HTS-XT accessory",
#          scan.mir.model.code_any_txt = "Bruker_Tensor_27.HTS.XT",
#          scan.mir.method.optics_any_txt = "",
#          scan.mir.method.preparation_any_txt = "",
#          scan.mir.license.title_ascii_txt = "CC-BY",
#          scan.mir.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/",
#          scan.mir.doi_idf_url = "https://doi.org/10.34725/DVN/QXCWP1",
#          scan.mir.contact.name_utf8_txt = "Vagen, Tor-Gunnar (World Agroforestry)",
#          scan.mir.contact.email_ietf_txt = "afsis.info@africasoils.net")

# Final export
afsis1.mir.export <- afsis1.mir

mir.exp.file = path(dir, "ossl_mir_v1.3")
readr::write_csv(afsis1.mir.export, str_c(mir.exp.file, ".csv.gz"))
nanoparquet::write_parquet(afsis1.mir.export, str_c(mir.exp.file, ".parquet"))
```

## Quality control

The final table must be joined as follows:

- MIR is used as first reference for pairing with soil data.
- Soil lab data are left joined to MIR. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test, include=TRUE, echo=TRUE, eval=TRUE}
afsis1.availability <- afsis1.mir.export %>%
  select(id.layer_local_c, scan_mir.600_abs) %>%
  left_join(afsis1.soildata, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Paired information
afsis1.availability %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Duplicates were dropped
afsis1.availability %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n(), .groups = 'drop') %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
afsis1.mir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "absorbance") %>%
  dplyr::mutate(wavenumber = gsub("scan_mir.|_abs", "", wavenumber)) %>%
  dplyr::mutate(wavenumber = as.numeric(wavenumber)) %>%
  ggplot(aes(x = wavenumber, y = absorbance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = c(600, 1200, 1800, 2400, 3000, 3600, 4000),
                     transform = "reverse") +
  labs(x = bquote("Wavenumber"~(cm^-1)), y = "Absorbance") +
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```

## References
