---
title: "Africa Soil Information Service Phase II (AFSIS2) prep"
author: 
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Ran Zhi
    email: jsafanelli@woodwellclimate.org
  - name: Tomislav Hengl
    email: tom.hengl@opengeohub.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for standardizing and importing the ICRAF Soil and Plant Spectroscopy Laboratory: Africa Soil Information Service Phase II, which includes the National 
Soil Information Systems: TanSIS (Tanzania), NiSIS (Nigeria) and GhanSIS (Ghana).

Website: [Soil Spectroscopy for Global Good](https://soilspectroscopy.org)  \
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## Original data

Site data, soil lab data, and Mid-Infrared Spectra (MIR) from the ICRAF Soil and Plant Spectroscopy Laboratory: Africa Soil Information Service Phase II.

The three datasets are available from <https://doi.org/10.34725/DVN/XUDGJY>, <https://doi.org/10.34725/DVN/WLAKR2> and <https://doi.org/10.34725/DVN/SPRSFN>. 

Data has been analyzed at the ICRAF Soil-Plant Spectral Diagnostics Laboratory, Nairobi, and the Rothamsted Research.

Coordinates for points are not provided and soil reference data is available only for a limited subset of points.

Further information about the datasets can be found in @Vagen2018 and @Hengl2021.

Coordinates for TanSIS points are available with a different coding system via <https://registry.opendata.aws/afsis/>.

Input datasets:  
- `glob2rx("*samples.csv")`: Separate csv files with site and soil analytes.\
- `afsis_mir_2013`: folder with MIR soil spectral data.\
- `Calibration_MPA_NIR.csv`: VNIR soil spectral data.\

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "data.table", "lubridate", "units", "readxl", "stringr",
              "googledrive", "googlesheets4", "tmap", "sf", "rnaturalearth",
              "skimr", "prospectr", "tictoc", "nanoparquet", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
gs4_auth(path = "~/secrets/soilcarbon-soilspec-845d0d3f2fbe.json")
options(scipen = 999)
```

Directory/folder path with original files (not uploaded to GitHub). 
```{r}
# dir = "./"
dir = "~/mnt-ossl-private/database/datasets/AFSIS2/"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
sample.files <- list.files(dir, pattern = glob2rx("*samples.csv"), full.names = T)

samples <- sample.files %>%
  purrr::map_dfr(fread, header = TRUE)

# Importing AfSIS 1 to see if some site info is available
afsis1.geo <- fread(paste0(dir, "/AfSIS1_georeferences.csv"))
tansis.geo <- fread(paste0(dir, "/AfSIS2_TanSIS_georeferences.csv"))

afsis1.sitedata <- afsis1.geo %>%
  select(SSN, Latitude, Longitude, Depth,
         Scientist, Cluster, Plot, Site, Country, Region, Cultivated) %>%
  rename(id.layer_local_c = SSN,
         longitude.point_wgs84_dd = Longitude,
         latitude.point_wgs84_dd = Latitude,
         surveyor.title_utf8_txt = Scientist,
         site.cluster_src_id = Cluster,
         site.plot_src_id = Plot,
         site.name_src_txt = Site,
         site.cultivated_src_logical = Cultivated,
         loc.country_src_txt = Country,
         loc.region_src_txt = Region) %>%
  mutate(observation.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2011-01-01"),
         observation.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2013-12-31")) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c),
         layer.upper.depth_usda_cm = ifelse(Depth == "top", 0, 20),
         layer.lower.depth_usda_cm = ifelse(Depth == "top", 20, 50),
         layer.sequence_usda_uint16 = ifelse(Depth == "top", 1, 2)) %>%
  select(-Depth)

tansis.sitedata <- tansis.geo %>%
  select(SSN, Latitude, Longitude, Depth,
         Scientist, Cluster, Plot, Site, Country, Region, Cultivated) %>%
  mutate(SSN = gsub("_|-", "", SSN)) %>%
  rename(id.layer_local_c = SSN,
         longitude.point_wgs84_dd = Longitude,
         latitude.point_wgs84_dd = Latitude,
         surveyor.title_utf8_txt = Scientist,
         site.cluster_src_id = Cluster,
         site.plot_src_id = Plot,
         site.name_src_txt = Site,
         site.cultivated_src_logical = Cultivated,
         loc.country_src_txt = Country,
         loc.region_src_txt = Region) %>%
  mutate(observation.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2018-01-01"),
         observation.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2018-12-31")) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  mutate(layer.upper.depth_usda_cm = ifelse(Depth == "top", 0, 20),
         layer.lower.depth_usda_cm = ifelse(Depth == "top", 20, 50),
         layer.sequence_usda_uint16 = ifelse(Depth == "top", 1, 2)) %>%
  select(-Depth)

# Removing unnecessary data
reference.files <- list.files(dir, pattern = glob2rx("*reference.csv"), full.names = T)

references <- reference.files %>%
  purrr::map_dfr(fread, header = TRUE) %>%
  mutate(SSN = str_to_upper(SSN))

reference.ids <- references %>%
  pull(SSN)

# Joining both datasets
afsis2.sitedata <- bind_rows(afsis1.sitedata, tansis.sitedata)

afsis2.sitedata <- samples %>%
  select(SSN, Country) %>%
  left_join(afsis2.sitedata, by = c("SSN" = "id.layer_local_c")) %>%
  mutate(SSN = str_to_upper(SSN)) %>%
  rename(id.layer_local_c = SSN,
         loc.country_src_txt2 = Country) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  filter(id.layer_local_c %in% reference.ids) %>%
  mutate(observation.date.begin_iso.8601_yyyy.mm.dd = lubridate::ymd("2018-01-01"),
         observation.date.end_iso.8601_yyyy.mm.dd = lubridate::ymd("2018-12-31")) %>%
  # mutate(id.project_ascii_txt = "TanSIS, NiSIS and GhanSIS (AfSIS-II) SSL",
  #        layer.texture_usda_txt = "",
  #        pedon.taxa_usda_txt = "",
  #        horizon.designation_usda_txt = "",
  #        longitude.county_wgs84_dd = NA,
  #        latitude.county_wgs84_dd = NA,
  #        location.point.error_any_m = 30,
  #        location.country_iso.3166_txt = "",
  #        observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
  #        observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
  #        surveyor.title_utf8_txt = case_when(is.na(Country) ~ NA_character_,
  #                                            Country == "Ghana" ~ "Teteh, Francis (CSIR)",
  #                                            Country == "Nigeria" ~ "Vincent, Aduramigba-Modupe (Obafemi Awolowo University)",
  #                                            Country == "Tanzania" ~ "Meliyo, Joel (TARI)",
  #                                            TRUE ~ NA_character_),
  #        surveyor.contact_ietf_email = "afsis.info@africasoils.net",
  #        surveyor.address_utf8_txt = "ICRAF, PO Box 30677, Nairobi, 00100, Kenya",
  #        dataset.title_utf8_txt = "TanSIS, NiSIS and GhanSIS (AfSIS-II) SSL",
  #        dataset.owner_utf8_txt = "the World Agroforestry Centre (ICRAF) and Rothamsted Research (RRES)",
  #        dataset.code_ascii_txt = "AFSIS2.SSL",
  #        dataset.address_idn_url = "https://www.isric.org/explore/ISRIC-collections",
  #        dataset.doi_idf_url = case_when(is.na(Country) ~ NA_character_,
  #                                        Country == "Ghana" ~ "https://doi.org/10.34725/DVN/SPRSFN",
  #                                        Country == "Nigeria" ~ "https://doi.org/10.34725/DVN/WLAKR2",
  #                                        Country == "Tanzania" ~ "https://doi.org/10.34725/DVN/XUDGJY",
  #                                        TRUE ~ NA_character_),
  #        dataset.license.title_ascii_txt = "CC0",
  #        dataset.license.address_idn_url = "https://creativecommons.org/publicdomain/zero/1.0/",
  #        dataset.contact.name_utf8_txt = "Winowiecki, Leigh Ann (ICRAF)",
  #        dataset.contact_ietf_email = "L.A.WINOWIECKI@cgiar.org") %>%
  mutate(loc.country_src_txt = ifelse(is.na(loc.country_src_txt),
                                       loc.country_src_txt2,
                                       loc.country_src_txt)) %>%
  select(-loc.country_src_txt2) %>%
  mutate(dataset.code_ascii_txt = "AFSIS2",
         id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .before = 1) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.exp.file = path(dir, "ossl_soilsite_v1.3")
readr::write_csv(afsis2.sitedata, str_c(site.exp.file, ".csv.gz"))
nanoparquet::write_parquet(afsis2.sitedata, str_c(site.exp.file, ".parquet"))
```

Plotting map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

ocean <- ne_download(scale = 110, type = "ocean", category = "physical", returnclass = "sf")

countries <- World[World$name %in% c("Ghana","Tanzania","Nigeria"), ]

tmap_mode("plot")
tm_shape(ocean) +
  tm_polygons(fill = "lightblue", col = NA) +
  tm_shape(World) +
  tm_polygons(fill = "#f0f0f0", fill_alpha = 0.5, col_alpha = 0.5) +
  tm_shape(countries) +
  tm_polygons(fill = "firebrick") +
  tm_crs("ESRI:54030") +
  tm_layout(frame = FALSE)
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template for scripted column standardization. Just run once for getting the original names of soil properties, descriptions, data types, and units. Then upload to Google Sheet for editing and manually defining the rules for integrating with the OSSL. Requires Google authentication. A copy of the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwritting, then rename online to download locally.**

```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables
variables.files = list.files(dir, glob2rx("0Variables Description*"),
                             full.names=TRUE, recursive = TRUE)
variables.files

# GhanSIS
readxl::excel_sheets(grep("GhanSIS", variables.files, value = T))
variables.names.GhanSIS <- readxl::read_xlsx(grep("GhanSIS", variables.files, value = T),sheet = 1)

# NiSIS
readxl::excel_sheets(grep("NiSIS", variables.files, value = T))
variables.names.NiSIS <- readxl::read_xlsx(grep("NiSIS", variables.files, value = T),sheet = 1)

# TanSIS
readxl::excel_sheets(grep("TanSIS", variables.files, value = T))
variables.names.TanSIS <- readxl::read_xlsx(grep("TanSIS", variables.files, value = T),sheet = 1)
names(variables.names.TanSIS)

# They all have the same soil properties
soillab.names <- variables.names.TanSIS %>%
  dplyr::mutate(table = "GhanSIS;NiSIS;TanSIS", .before = 1) %>%
  dplyr::rename(original_name = `variable name`, original_description = `variable description`,
                units = Units) %>%
  dplyr::mutate(comment = paste0(`method used`, "; ", `instrument used for analysis`)) %>%
  dplyr::select(table, original_name, original_description, units, comment) %>%
  dplyr::mutate(import = '', ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, paste0(getwd(), "/afsis2_soillab_names.csv"))

# Uploading to google sheet

# FACT CIN folder. Get ID for soildata importing table
googledrive::drive_ls(as_id("0AHDIWmLAj40_Uk9PVA"))

OSSL.soildata.importing <- "19LeILz9AEnKVK7GK0ZbK3CCr2RfeP-gSWn5VpY8ETVM"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names
upload <- dplyr::as_tibble(soillab.names)

# Uploading
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "AFSIS2")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the edited standardization template is saved to this dataset folder.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "AFSIS2")

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:
```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F) %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "ossl_")))

knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:
```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
reference.files <- list.files(dir, pattern = glob2rx("*reference.csv"), full.names = T)

afsis2.reference <- reference.files %>%
  purrr::map_dfr(fread, header = TRUE)

# Harmonization of names and units
analytes.old.names <- transvalues %>%
  filter(table == "GhanSIS;NiSIS;TanSIS") %>%
  pull(original_name)

analytes.new.names <- transvalues %>%
  filter(table == "GhanSIS;NiSIS;TanSIS") %>%
  pull(ossl_name)

# Selecting and renaming
afsis2.soildata <- afsis2.reference %>%
  rename(id.layer_local_c = SSN) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  select(id.layer_local_c, all_of(analytes.old.names)) %>%
  rename_with(~analytes.new.names, all_of(analytes.old.names)) %>%
  as.data.frame()

# Removing duplicates
# afsis2.soildata %>%
#   group_by(id.layer_local_c) %>%
#   summarise(repeats = n()) %>%
#   group_by(repeats) %>%
#   summarise(count = n())

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "GhanSIS;NiSIS;TanSIS") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(afsis2.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
afsis2.soildata.trans <- transform_values(df = afsis2.soildata,
                                          out.name = names(afsis2.soildata),
                                          in.name = names(afsis2.soildata),
                                          fun.lst = functions.list)

# Final soillab data
afsis2.soildata <- afsis2.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
afsis2.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.exp.file = path(dir, "ossl_soillab_v1.3")
readr::write_csv(afsis2.soildata, str_c(soillab.exp.file, ".csv.gz"))
nanoparquet::write_parquet(afsis2.soildata, str_c(soillab.exp.file, ".parquet"))
```

Soil lab data summary.
```{r soil_lab_summary, include=TRUE, echo=TRUE, eval=TRUE}
afsis2.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### Mid-infrared spectra

```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
# Floating wavenumbers
mir.files <- list.files(dir, pattern = glob2rx("*_ZnSe_*.csv$"), full.names = T)

mir.scans <- mir.files %>%
  purrr::map_dfr(fread, header = TRUE, colClasses = 'character')

# # Some prefixes are either wrong or shuffled for anonimity
# mir.scans %>%
#   mutate(prefix = str_sub(SSN, 1, 7)) %>%
#   distinct(prefix)
# 
# mir.scans.ids <- mir.scans %>%
#   select(SSN) %>%
#   mutate(SSN = str_to_upper(SSN)) %>%
#   mutate(SSN = gsub("_|-", "", SSN)) %>%
#   mutate(SSN = gsub(" ", "", SSN)) %>%
#   mutate(SSN = gsub("\\b(ASAG)", "TANSIS", SSN)) %>%
#   mutate(SSN = gsub("\\b(ANSIS)", "TANSIS", SSN)) %>%
#   mutate(SSN = gsub("\\b(ANSAG)", "TANSIS", SSN)) %>%
#   mutate(SSN = gsub("\\b(TASAG)", "TANSIS", SSN)) %>%
#   mutate(SSN = gsub("\\b(TANSAG)", "TANSIS", SSN)) %>%
#   mutate(prefix = str_sub(SSN, 1, -9),
#          sufix = str_sub(SSN, -8, -1))
# 
# mir.scans.ids %>%
#   mutate(prefix = str_sub(SSN, 1, 6)) %>%
#   distinct(prefix)
# 
# soildata.ids <- afsis2.soildata %>%
#   select(id.layer_local_c) %>%
#   mutate(id.layer_local_c = str_to_upper(id.layer_local_c)) %>%
#   mutate(prefix = str_sub(id.layer_local_c, 1, -9),
#          sufix = str_sub(id.layer_local_c, -8, -1))
# 
# soilsite.ids <- afsis2.sitedata %>%
#   select(id.layer_local_c) %>%
#   mutate(id.layer_local_c = str_to_upper(id.layer_local_c)) %>%
#   mutate(prefix = str_sub(id.layer_local_c, 1, -9),
#          sufix = str_sub(id.layer_local_c, -8, -1))
# 
# table(soildata.ids$sufix %in% mir.scans.ids$sufix)
# table(soilsite.ids$sufix %in% mir.scans.ids$sufix)
# table(soildata.ids$sufix %in% soilsite.ids$sufix)
# table(soildata.ids$id.layer_local_c %in% soilsite.ids$id.layer_local_c)

mir.scans <- mir.scans %>%
  mutate(SSN = str_to_upper(SSN))

old.names <- names(mir.scans)
new.names <- gsub("X", "", old.names)

afsis2.mir <- mir.scans %>%
  rename_with(~new.names, all_of(old.names)) %>%
  rename(id.layer_local_c = SSN) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  mutate(across(-id.layer_local_c, ~as.numeric(.x)))

# Need to resample spectra
old.wavenumber <- na.omit(as.numeric(names(afsis2.mir)))
new.wavenumbers <- rev(seq(600, 4000, by = 2))

afsis2.mir <- afsis2.mir %>%
  select(-id.layer_local_c) %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavenumber,
                      new.wav = new.wavenumbers, interpol = "spline") %>%
  as_tibble() %>%
  bind_cols({afsis2.mir %>%
      select(id.layer_local_c)}, .) %>%
  select(id.layer_local_c, as.character(rev(new.wavenumbers)))

# Gaps
scans.na.gaps <- afsis2.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({afsis2.mir %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
scans.extreme.neg <- afsis2.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < -1, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({afsis2.mir %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
scans.extreme.pos <- afsis2.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({afsis2.mir %>% select(id.layer_local_c)}, .)

# Consistency summary - problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
old.wavenumbers <- seq(600, 4000, by = 2)
new.wavenumbers <- paste0("scan_mir.", old.wavenumbers, "_abs")

afsis2.mir <- afsis2.mir %>%
  rename_with(~new.wavenumbers, all_of(as.character(old.wavenumbers)))

# # Preparing metadata
# afsis2.mir.metadata <- afsis2.mir %>%
#   select(id.layer_local_c) %>%
#   mutate(id.scan_local_c = id.layer_local_c) %>%
#   mutate(scan.mir.date.begin_iso.8601_yyyy.mm.dd = ymd("2019-01-01"),
#          scan.mir.date.end_iso.8601_yyyy.mm.dd = ymd("2019-12-31"),
#          scan.mir.model.name_utf8_txt = "Bruker Alpha 1 ZnSe",
#          scan.mir.model.code_any_txt = "Bruker_Alpha_1_ZnSe",
#          scan.mir.method.optics_any_txt = "",
#          scan.mir.method.preparation_any_txt = "",
#          scan.mir.license.title_ascii_txt = "CC0",
#          scan.mir.license.address_idn_url = "https://creativecommons.org/publicdomain/zero/1.0/",
#          scan.mir.doi_idf_url = "https://doi.org/10.34725/DVN/XUDGJY",
#          scan.mir.contact.name_utf8_txt = "Winowiecki, Leigh Ann (ICRAF)",
#          scan.mir.contact.email_ietf_txt = "L.A.WINOWIECKI@cgiar.org")

# # Final preparation
# afsis2.mir.export <- afsis2.mir.metadata %>%
#   left_join(afsis2.mir, by = "id.layer_local_c") %>%
#   mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
afsis2.mir.export <- afsis2.mir

mir.exp.file = path(dir, "ossl_mir_v1.3")
readr::write_csv(afsis2.mir.export, str_c(mir.exp.file, ".csv.gz"))
nanoparquet::write_parquet(afsis2.mir.export, str_c(mir.exp.file, ".parquet"))
```

## Quality control

The final table must be joined as follows:

- MIR is used as first reference for pairing with soil data.
- Soil lab data are left joined to MIR. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
afsis2.availability <- afsis2.mir.export %>%
  select(id.layer_local_c, scan_mir.800_abs) %>%
  left_join(afsis2.soildata, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from afsis2
afsis2.availability %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Duplicates are dropped
afsis2.availability %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
afsis2.mir %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "absorbance") %>%
  dplyr::mutate(wavenumber = gsub("scan_mir.|_abs", "", wavenumber)) %>%
  dplyr::mutate(wavenumber = as.numeric(wavenumber)) %>%
  ggplot(aes(x = wavenumber, y = absorbance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = c(600, 1200, 1800, 2400, 3000, 3600, 4000),
                     transform = "reverse") +
  labs(x = bquote("Wavenumber"~(cm^-1)), y = "Absorbance") +
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```

## References
