---
title: "GEMAS dataset preparation for the OSSL"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
  - name: Federal Institute for Geosciences and Natural Resources 
    email: gemas@bgr.de
  - name: José Martín Soriano Disla
    email: martin.soriano@cetenma.es
  - name: Leslie J. Janik
    email: les.janik@csiro.au
  - name: Clemens Reimann (former lead, retired), Geological Survey of Norway (NGU)
    email: clemensreimann@yahoo.co.uk
  - name: Anna Ladenberger, Geological Survey of Sweden (SGU)
    email: anna.ladenberger@sgu.se
  - name: Philippe Negrel, French Geological Survey (BRGM)
    email: p.negrel@brgm.fr
format: 
  gfm:
    toc: true
    toc-depth: 2
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Geochemical Mapping of Agricultural and Grazing Land Soil (GEMAS) dataset into the Open Soil Spectral Library.

Website: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The GEMAS original data

Site data, Soil lab data, and Mid-Infrared Spectra (MIR) from the Geochemical Mapping of Agricultural and Grazing Land Soil (GEMAS). Further information of the dataset can be found in detail at https://gemas.eurogeosurveys.org/, @reimann_chemistry_2014, @reimann_chemistry_2014-1, and @soriano-disla_use_2013.

Original files:\
- `GEMAS_MIRcompiled.csv`: csv file with site information, soil information, and spectral data.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "prospectr", "units", "readxl", "stringr",
              "olctools", "openssl", "tmap", "sf", "skimr", "lubridate",
              "googledrive", "googlesheets4", "data.table", "doMC", "tictoc",
              "qs", "dplyr", "tidyr", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r}
dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/GEMAS"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
gemas.metadata <- fread(path(dir, "GEMAS_MIRcompiled.csv"), header = T)

gemas.sitedata <- gemas.metadata %>%
  select(aowx_file, sample_id, COUNTRY, UHDICM, LHDICM, TYPE, XCOO, YCOO) %>%
  rename(id.layer_local_c = aowx_file,
         id.sample_local_c = sample_id,
         longitude.point_wgs84_dd = XCOO,
         latitude.point_wgs84_dd = YCOO,
         layer.upper.depth_usda_cm = UHDICM,
         layer.lower.depth_usda_cm = LHDICM) %>%
  mutate(across(c(id.layer_local_c, id.sample_local_c), as.character)) %>%
  mutate(layer.sequence_usda_uint16 = ifelse(layer.upper.depth_usda_cm == 0, 1, 2)) %>%
  mutate(id.project_ascii_txt = "GEMAS Soil Spectral Library",
         dataset.code_ascii_txt = "GEMAS.MIR",
         observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
         observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
         surveyor.title_utf8_txt = "Clemens Reimann, Manfred Birke, Alecos Demetriades, and Ilse Schoeters, Rio Tinto et.al ",
         surveyor.contact_ietf_email = " Clemens.Reimann@ngu.no, Manfred.Birke@bgr.de, ademetriades@igme.gr, and Ilse.SCHOETERS@riotinto.com",
         dataset.title_utf8_txt = "GEMAS Soil Spectral Library",
         dataset.owner_utf8_txt = "EuroGeoSurveys Geochemistry Expert Group and Eurometaux",
         dataset.address_idn_url = "https://gemas.eurogeosurveys.org/GEMAS.htm",
         dataset.doi_idf_url = "https://data.europa.eu/data/datasets/20607a6b-f412-4777-9d43-67d734677d57?locale=en",
         dataset.license.title_ascii_txt = "CC-BY 4.0",
         dataset.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/legalcode",
         dataset.contact.name_utf8_txt = "Federal Institute for Geosciences and Natural Resources",
         dataset.contact_ietf_email = "gemas@bgr.de") %>%
  mutate(id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         id.location_olc_txt = olctools::encode_olc(latitude.point_wgs84_dd, longitude.point_wgs84_dd, 10),
         .after = id.project_ascii_txt) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.qs = path(dir, "ossl_soilsite_v2.0.qs")
qs::qsave(gemas.sitedata, site.qs, preset = "high")
```

Plotting sites map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

points <- gemas.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")

tm_shape(World) +
  tm_polygons(fill = '#f0f0f0', col_alpha = 0.2) +
tm_shape(points) +
  tm_dots(fill = "COUNTRY", 
          fill.scale = tm_scale_categorical(values = "brewer.set1", n.max = 40)) +
  tm_layout(legend.outside = TRUE)

```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.


```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- gemas.metadata %>%
  select(aowx_file, sample_id, silt, clay, As, Cd, Cu, Pb, Zn, C_tot, TOC,
         CEC, pH_CaCl2, LOI) %>%
  names(.) %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'GEMAS_MIRcompiled.csv', .before = 1) %>%
  dplyr::mutate(import = '', original_unit = '',  original_method = '', comment = '', ossl_abbrev = '', ossl_method = '', 
                ossl_unit = '', ossl_convert = '', ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the output file is saved to this folder for archiving purposes.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "GEMAS") %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
# Firstly, average the replicates
gemas.averaged <- gemas.metadata %>%
  group_by(sample_id, COUNTRY, UHDICM, LHDICM, C_ID, TYPE, XCOO, YCOO, XLAEA, YLAEA, ALT) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)), .groups = "drop")

gemas.reference <- gemas.averaged

# Harmonization of names and units
analytes.old.names <- transvalues %>%
  filter(table == "GEMAS_MIRcompiled.csv") %>%
  pull(original_name)

analytes.new.names <- transvalues %>%
  filter(table == "GEMAS_MIRcompiled.csv") %>%
  pull(ossl_name)

# Selecting and renaming
analytes.old.names.clean <- analytes.old.names[analytes.old.names != "sample_id"]
analytes.new.names.clean <- analytes.new.names[analytes.old.names != "sample_id"]

gemas.soildata <- gemas.reference %>%
  rename(id.layer_local_c = sample_id) %>%
  select(id.layer_local_c, all_of(analytes.old.names.clean)) %>%
  rename_with(~analytes.new.names.clean, all_of(analytes.old.names.clean)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "GEMAS_MIRcompiled.csv") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(gemas.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
gemas.soildata.trans <- transform_values(df = gemas.soildata,
                                       out.name = names(gemas.soildata),
                                       in.name = names(gemas.soildata),
                                       fun.lst = functions.list)

# Final soillab data
gemas.soildata <- gemas.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
gemas.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.qs = path(dir, "ossl_soillab_v2.0.qs")
qs::qsave(gemas.soildata, soillab.qs, preset = "high")
```

Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
gemas.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```


### Mid-infrared spectra (MIR)
```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
gemas.mir.proc <- gemas.averaged %>%
  rename(id.layer_local_c = sample_id) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# Need to resample spectra
old.wavenumber <- names(gemas.mir.proc)[grep("^[0-9]+$", names(gemas.mir.proc))]
old.wavenumber.num <- as.numeric(old.wavenumber)

new.wavenumber <- seq(600, 4000, by = 2)

gemas.mir.matrix <- gemas.mir.proc %>%
  select(all_of(old.wavenumber)) %>%
  as.matrix()

gemas.mir.resampled <- prospectr::resample(
  X = gemas.mir.matrix, 
  wav = old.wavenumber.num, 
  new.wav = new.wavenumber, 
  interpol = "spline"
) %>%
  as_tibble()

gemas.mir.proc <- bind_cols(
  gemas.mir.proc %>% select(id.layer_local_c), 
  gemas.mir.resampled
)

# Gaps

mir.na.gaps <- gemas.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({gemas.mir.proc %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
mir.extreme.neg <- gemas.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({gemas.mir.proc %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
mir.extreme.pos <- gemas.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({gemas.mir.proc %>% select(id.layer_local_c)}, .)


# Consistency summary - problematic scans
mir.summary <- mir.na.gaps %>%
  left_join(mir.extreme.neg, by = "id.layer_local_c") %>%
  left_join(mir.extreme.pos, by = "id.layer_local_c")

mir.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
final.mir.names <- paste0("scan_mir.", new.wavenumber, "_ref")
gemas.mir.proc <- gemas.mir.proc %>%
  rename_with(~final.mir.names, as.character(new.wavenumber))

# Preparing metadata
gemas.mir.metadata <- gemas.mir.proc %>%
  select(id.layer_local_c) %>%
  mutate(id.scan_local_c = id.layer_local_c,
         scan.mir.date.begin_iso.8601_yyyy = ymd("2008-01-01"), 
         scan.mir.date.end_iso.8601_yyyy = ymd("2009-3-30"), 
         scan.mir.model.name_utf8_txt = "Perkin–Elmer Spectrum-One™ Fourier-transform infrared spectrometer (Perkin Elmer Inc., Mass. USA)", 
         scan.mir.license.title_ascii_txt = "CC-BY 4.0",
         scan.mir.method.optics_any_txt = "",
         scan.mir.method.preparation_any_txt = "All soil samples were air dried, sieved to <2 mm using a nylon screen, homogenised and finally split into sub-samples (10 splits)",
         scan.mir.doi_idf_url = "https://gemas.eurogeosurveys.org/GEMAS.htm",
         scan.mir.contact.name_utf8_txt = "Clemens Reimann",
         scan.mir.contact.email_ietf_txt = "Clemens.Reimann@ngu.no")

# Final preparation
gemas.mir.export <- gemas.mir.metadata %>%
  left_join(gemas.mir.proc, by = "id.layer_local_c") %>%
  mutate(across(starts_with("id."), as.character))

# Saving version to dataset root dir
soilmir.qs = path(dir, "ossl_mir_v2.0.qs")
qs::qsave(gemas.mir.export, soilmir.qs, preset = "high")
```

### Quality control 

The final table must be joined as follows:

- MIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_mir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
gemas.availability <- gemas.mir.export %>%
  select(id.layer_local_c, scan_mir.600_ref) %>%
  left_join({gemas.sitedata %>%
      select(id.layer_local_c, layer.lower.depth_usda_cm)}, by = "id.layer_local_c") %>%
  left_join({gemas.soildata %>%
      select(id.layer_local_c, any_of("clay_w.pct"))}, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from besb
gemas.info.summary <- gemas.availability %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())
print("Availability Summary:")
print(gemas.info.summary)

# Repeats check - Duplicates are dropped
gemas.availability %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

Soil analytical data summary for MIR. Note: many scans could not be linked with the wetchem.
```{r summary_mir, include=TRUE, echo=TRUE, eval=TRUE}
gemas.soildata %>%
  filter(id.layer_local_c %in% gemas.mir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
gemas.mir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "absorbance") %>%
  dplyr::mutate(wavenumber = as.numeric(gsub("scan_mir.|_ref", "", wavenumber))) %>%
  ggplot(aes(x = wavenumber, y = absorbance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(600, 4000, by = 400),
                     transform = "reverse") +
  labs(title = "MIR Spectra (100 random scans)",
       x = bquote("Wavenumber"~(cm^-1)),
       y = "Absorbance")+
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```
## References

