---
title: "BESB dataset preparation for the OSSL"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
date: last-modified
date-format: "--- DD MMMM, YYYY."
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Brazilian Soil Spectral
Library (BSSL) dataset into the Open Soil Spectral Library.

Project: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The BSSL original data

Site data, Soil lab data, Mid-Infrared Spectra (MIR), and Visible,
Near-Infrared and Shortwave-Infrared (Vis-NIR-SWIR) from the Brazilian
Soil Spectral Library (BSSL). Further information of the dataset can be
found in detail at @dematte_brazilian_2019 and @dematte_brazilian_2022.

Original files:\
- `BSSL_DB_V002.xlsx`: xlsx file with site information, soil information, and spectral data.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "prospectr", "units", "readxl", "stringr",
              "olctools", "openssl", "tmap", "sf", "skimr", "lubridate",
              "googledrive", "googlesheets4", "data.table", "doMC", "tictoc",
              "qs", "dplyr", "tidyr", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r}
dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/BESB"
# dir = "~/projects/mnt-ossl/import/dataset/BESB"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
besb.metadata <- read_excel(path(dir, "BSSL_DB_V002.xlsx"),
                            sheet = "BSSL_Soil_Attributes_Dataset")

besb.sitedata <- besb.metadata %>%
  select(ID_Unique, Owner_code, Vis_NIR_SWIR_availability, MIR_availability, Depth_cm,
         Lat, Long, Region, Vegetation, Bioma, Geology) %>%
  rename(id.layer_local_c = ID_Unique,
         longitude.point_wgs84_dd = Long,
         latitude.point_wgs84_dd = Lat) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  separate(Depth_cm, into = c("layer.upper.depth_usda_cm", "layer.lower.depth_usda_cm"), sep = "-") %>%
  mutate(across(starts_with("layer."), as.numeric)) %>%
  mutate(layer.sequence_usda_uint16 = ifelse(layer.upper.depth_usda_cm == 0, 1, 2)) %>%
  mutate(id.project_ascii_txt = "Brazilian Soil Spectral Library",
         dataset.code_ascii_txt = "Brazilian.SSL",
         observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
         observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
         surveyor.title_utf8_txt = "Luiz de Queiroz College of Agriculture, University of São Paulo, Piracicaba, São Paulo, Brazil",
         surveyor.contact_ietf_email = "jamdemat@usp.br",
         surveyor.address_utf8_txt = "Department of Soil Science, Luiz de Queiroz College of Agriculture (ESALQ), University of São Paulo (USP), Ave. Pádua Dias 11, Cx. Postal 9, 13418-900, Piracicaba, São Paulo, Brazil",
         dataset.title_utf8_txt = "Brazilian Soil Spectral Library",
         dataset.owner_utf8_txt = "Luiz de Queiroz College of Agriculture, University of São Paulo",
         
         dataset.address_idn_url = "https://bibliotecaespectral.wixsite.com/english",
         dataset.doi_idf_url = "https://doi.org/10.5281/zenodo.8361419",
         dataset.license.title_ascii_txt = "CC-BY",
         dataset.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/legalcode",
         dataset.contact.name_utf8_txt = "José A.M. Demattê",
         dataset.contact_ietf_email = "jamdemat@usp.br") %>%
  mutate(id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         id.location_olc_txt = olctools::encode_olc(latitude.point_wgs84_dd, longitude.point_wgs84_dd, 10),
         .after = id.project_ascii_txt) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.qs = path(dir, "ossl_soilsite_v2.0.qs")
qs::qsave(besb.sitedata, site.qs, preset = "high")
```

Plotting sites map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

points <- besb.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")

tm_shape(World) +
  tm_polygons('#f0f0f0f0', border.alpha = 0.2) +
  tm_shape(points) +
  tm_dots()
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwritting, then rename online to download locally.**

```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- besb.metadata %>%
  select(ID_Unique, Sand_gkg, Clay_gkg, SOM_gkg, pH_H2O, 
         Ca_mmolkg, Mg_mmolkg, Na_mmolkg, CEC_Ph7_mmolkg) %>%
  names(.) %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'BSSL_DB_V002.xlsx', .before = 1) %>%
  dplyr::mutate(import = '', comment = '', ossl_abbrev = '', ossl_method = '', ossl_unit = '',
                ossl_convert = '', ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the output file is saved to this folder for archiving purposes.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "BESB") %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
besb.reference <- besb.metadata

# Harmonization of names and units
analytes.old.names <- transvalues %>%
  filter(table == "BSSL_DB_V002.xlsx") %>%
  pull(original_name)

analytes.new.names <- transvalues %>%
  filter(table == "BSSL_DB_V002.xlsx") %>%
  pull(ossl_name)

# Selecting and renaming
if("SOM_gkg" %in% names(besb.reference)) {
  besb.reference <- besb.reference %>% rename(SOC_gkg = SOM_gkg)
}

analytes.old.names.clean <- analytes.old.names[analytes.old.names != "ID_Unique"]
analytes.new.names.clean <- analytes.new.names[analytes.old.names != "ID_Unique"]

besb.soildata <- besb.reference %>%
  rename(id.layer_local_c = ID_Unique) %>%
  select(id.layer_local_c, all_of(analytes.old.names.clean)) %>%
  rename_with(~analytes.new.names.clean, all_of(analytes.old.names.clean)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "BSSL_DB_V002.xlsx") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(besb.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
besb.soildata.trans <- transform_values(df = besb.soildata,
                                       out.name = names(besb.soildata),
                                       in.name = names(besb.soildata),
                                       fun.lst = functions.list)

# Final soillab data
besb.soildata <- besb.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
besb.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.qs = path(dir, "ossl_soillab_v2.0.qs")
qs::qsave(besb.soildata, soillab.qs, preset = "high")
```

Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
besb.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### Vis-NIR spectra

```{r vis-nir, include=TRUE, echo=TRUE, eval=TRUE}
# Floating wavenumbers

besb.visnir <- read_excel(path(dir, "BSSL_DB_V002.xlsx"),
                            sheet = "Vis_NIR_SWIR_Dataset")

# Renaming
besb.visnir.proc <- besb.visnir %>%
  rename(id.layer_local_c = ID_Unique) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# Need to resample spectra
old.wavelengths <- na.omit(as.numeric(names(besb.visnir.proc)))
new.wavelengths <- seq(350, 2500, by = 2)

besb.visnir.proc <- besb.visnir.proc %>%
  select(-id.layer_local_c, -Owner_code) %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavelengths, new.wav = new.wavelengths, interpol = "spline") %>%
  as_tibble() %>%
  bind_cols({besb.visnir.proc %>%
      select(id.layer_local_c)}, .) %>%
  select(id.layer_local_c, as.character(new.wavelengths))

# Gaps

scans.na.gaps <- besb.visnir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({besb.visnir.proc %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
scans.extreme.neg <- besb.visnir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({besb.visnir.proc %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
scans.extreme.pos <- besb.visnir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 1.5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({besb.visnir.proc %>% select(id.layer_local_c)}, .)

# Consistency summary - problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
final.visnir.names <- paste0("scan_visnir.", new.wavelengths, "_ref")
besb.visnir.proc <- besb.visnir.proc %>%
  rename_with(~final.visnir.names, as.character(new.wavelengths))

# Preparing metadata
besb.visnir.metadata <- besb.visnir.proc %>%
  select(id.layer_local_c) %>%
  mutate(id.scan_local_c = id.layer_local_c,
         scan.visnir.date.begin_iso.8601_yyyy = ymd("1995-01-01"), 
         scan.visnir.date.begin_iso.8601_yyyy = ymd("2019-12-30"), 
         scan.visnir.model.name_utf8_txt = "ASD FieldSpec 3 spectroradiometer", 
         scan.visnir.license.title_ascii_txt = "CC-BY",
         scan.visnir.method.optics_any_txt = "",
         scan.visnir.method.preparation_any_txt = "Drying soil samples in an oven at 45°C for 48 hours, grinding them, and sieving them through a 2mm mesh.",
         scan.visnir.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/legalcode",
         scan.visnir.doi_idf_url = "https://doi.org/10.5281/zenodo.8361419",
         scan.visnir.contact.name_utf8_txt = "José A.M. Demattê",
         scan.visnir.contact.email_ietf_txt = "jamdemat@usp.br")


# Final preparation
besb.visnir.export <- besb.visnir.metadata %>%
  left_join(besb.visnir.proc, by = "id.layer_local_c") %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
soilvisnir.qs = path(dir, "ossl_visnir_v2.0.qs")
qs::qsave(besb.visnir.export, soilvisnir.qs, preset = "high")
```

### Quality control for Vis-NIR

The final table must be joined as follows:

- Vis-NIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_visnir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
besb.availability <- besb.visnir.export %>%
  select(id.layer_local_c, scan_visnir.600_ref) %>%
  left_join({besb.sitedata %>%
      select(id.layer_local_c, layer.lower.depth_usda_cm)}, by = "id.layer_local_c") %>%
  left_join({besb.soildata %>%
      select(id.layer_local_c, sand.tot_usda.c405_w.pct)}, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from besb
besb.availability %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Duplicates are dropped
besb.availability %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

Soil analytical data summary for Vis-NIR. Note: many scans could not be linked with the wetchem.
```{r summary_visnir, include=TRUE, echo=TRUE, eval=TRUE}
besb.soildata %>%
  filter(id.layer_local_c %in% besb.visnir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

Vis-NIR spectral visualization (100 random spectra):
```{r visnir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
besb.visnir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_visnir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavelength", values_to = "reflectance") %>%
  dplyr::mutate(wavelength = gsub("scan_visnir.|_ref", "", wavelength)) %>%
  dplyr::mutate(wavelength = as.numeric(wavelength)) %>%
  ggplot(aes(x = wavelength, y = reflectance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(350, 2500, by = 250))+
  labs(title = "Vis-NIR Spectra (100 random scans)",
       x = "Wavelength (nm)",
       y = "Reflectance")+
  theme_light()
```


### Mid-infrared spectra (MIR)
```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
besb.mir <- read_excel(path(dir, "BSSL_DB_V002.xlsx"),
                            sheet = "BSSL_MIR_Dataset")

# Renaming
besb.mir.proc <- besb.mir %>%
  rename(id.layer_local_c = ID_Unique) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# Need to resample spectra
old.wavenumber <- na.omit(as.numeric(names(besb.mir.proc)))
new.wavenumber <- seq(600, 4000, by = 2)

besb.mir.proc <- besb.mir.proc %>%
  select(-id.layer_local_c, -Owner_code) %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavenumber, new.wav = new.wavenumber, interpol = "spline") %>%
  as_tibble() %>%
  bind_cols({besb.mir.proc %>%
      select(id.layer_local_c)}, .) %>%
  select(id.layer_local_c, as.character(new.wavenumber))

# Gaps

mir.na.gaps <- besb.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({besb.mir.proc %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
mir.extreme.neg <- besb.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({besb.mir.proc %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
mir.extreme.pos <- besb.mir.proc %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({besb.mir.proc %>% select(id.layer_local_c)}, .)

# Consistency summary - problematic scans
mir.summary <- mir.na.gaps %>%
  left_join(mir.extreme.neg, by = "id.layer_local_c") %>%
  left_join(mir.extreme.pos, by = "id.layer_local_c")

mir.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
final.mir.names <- paste0("scan_mir.", new.wavenumber, "_ref")
besb.mir.proc <- besb.mir.proc %>%
  rename_with(~final.mir.names, as.character(new.wavenumber))

# Preparing metadata
besb.mir.metadata <- besb.mir.proc %>%
  select(id.layer_local_c) %>%
  mutate(id.scan_local_c = id.layer_local_c,
         scan.mir.date.begin_iso.8601_yyyy = ymd("1995-01-01"), 
         scan.mir.date.begin_iso.8601_yyyy = ymd("2019-12-30"), 
         scan.mir.model.name_utf8_txt = "Fourier Transform Infrared (FT-IR) alpha spectroradiometer", 
         scan.mir.license.title_ascii_txt = "CC-BY",
         scan.mir.method.optics_any_txt = "",
         scan.mir.method.preparation_any_txt = "Milled the soil fraction smaller than 2 mm, sieved it to 0.149 mm.",
         scan.mir.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/legalcode",
         scan.mir.doi_idf_url = "https://doi.org/10.5281/zenodo.8361419",
         scan.mir.contact.name_utf8_txt = "José A.M. Demattê",
         scan.mir.contact.email_ietf_txt = "jamdemat@usp.br")


# Final preparation
besb.mir.export <- besb.mir.metadata %>%
  left_join(besb.mir.proc, by = "id.layer_local_c") %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
soilmir.qs = path(dir, "ossl_mir_v2.0.qs")
qs::qsave(besb.mir.export, soilmir.qs, preset = "high")
```

### Quality control for MIR

The final table must be joined as follows:

- MIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_mir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
besb.availability2 <- besb.mir.export %>%
  select(id.layer_local_c, scan_mir.600_ref) %>%
  left_join({besb.sitedata %>%
      select(id.layer_local_c, layer.lower.depth_usda_cm)}, by = "id.layer_local_c") %>%
  left_join({besb.soildata %>%
      select(id.layer_local_c, sand.tot_usda.c405_w.pct)}, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from besb
besb.availability2 %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Duplicates are dropped
besb.availability2 %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

Soil analytical data summary for MIR. Note: many scans could not be linked with the wetchem.
```{r summary_mir, include=TRUE, echo=TRUE, eval=TRUE}
besb.soildata %>%
  filter(id.layer_local_c %in% besb.mir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
besb.mir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "reflectance") %>%
  dplyr::mutate(wavenumber = gsub("scan_mir.|_ref", "", wavenumber)) %>%
  dplyr::mutate(wavenumber = as.numeric(wavenumber)) %>%
  ggplot(aes(x = wavenumber, y = reflectance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = c(600, 1200, 1800, 2400, 3000, 3600, 4000),
                     transform = "reverse") +
  labs(title = "MIR Spectra (100 random scans)",
       x = bquote("Wavenumber"~(cm^-1)),
       y = "Reflectance")+
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```
## References

