---
title: "Spectra from two Brazilian farms"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the spectra from two Brazilian local farms into the Open Soil Spectral Library.

Website: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The Tropical farm original data

Site data, Soil lab data, and Visible Near-Infrared (Vis-NIR) data from Brazilian agricultural areas. Further information of the dataset can be found in detail at @Tavares2022.

Original files:\
- `soil fertility data.xlsx`: xlsx file with site information and soil lab information. 
- `VNIR data.xlsx`: xlsx file with Vis-NIR spectral data.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "prospectr", "units", "readxl", "stringr",
              "olctools", "openssl", "tmap", "sf", "skimr", "lubridate",
              "googledrive", "googlesheets4", "data.table", "doMC", "tictoc",
              "qs", "dplyr", "tidyr", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r}
dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/TropicalFarm"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
tropical.metadata <- read_excel(path(dir, "soil fertility data.xlsx"))

tropical.sitedata <- tropical.metadata %>%
  select(ID, Field) %>%
  rename(id.layer_local_c = ID) %>%
  mutate(layer.upper.depth_usda_cm = 0,
         layer.lower.depth_usda_cm = 20) %>%
  mutate(id.project_ascii_txt = "Spectral data of tropical soils using dry-chemistry techniques",
         dataset.code_ascii_txt = "Tropical_Farm.SSL",
         observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
         observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
         dataset.owner_utf8_txt = "Universidade de Sao Paulo",
         dataset.doi_idf_url = "https://data.mendeley.com/datasets/88c5kvmgbf/1",
         dataset.license.title_ascii_txt = "CC-BY 4.0",
         dataset.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/",
         dataset.contact.name_utf8_txt = "Tiago Rodrigues Tavares",
         dataset.contact_ietf_email = "tiagosrt@usp.br") %>%
  mutate(id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .after = id.project_ascii_txt) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.qs = path(dir, "ossl_soilsite_v2.0.qs")
qs::qsave(tropical.sitedata, site.qs, preset = "high")
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.


```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- tropical.metadata %>%
  select(ID, Clay, OM, CEC, pH, V, exP, exK, exCa, exMg) %>%
  rename(id.layer_local_c = ID) %>%
  names() %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'soil fertility data.xlsx', .before = 1) %>%
  dplyr::mutate(import = '', original_unit = '', original_method = '', comment = '', ossl_abbrev = '', ossl_method = '', 
                ossl_unit = '', ossl_convert = '', ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the output file is saved to this folder for archiving purposes.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "TropicalFarm") %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
tropical.reference <- tropical.metadata

# Harmonization of names and units
valid_mappings <- transvalues %>%
  filter(table == "soil fertility data.xlsx") %>%
  filter(!is.na(ossl_name) & ossl_name != "") 

analytes.old.names <- valid_mappings %>% pull(original_name)
analytes.new.names <- valid_mappings %>% pull(ossl_name)

tropical.soildata <- tropical.reference %>%
  rename(id.layer_local_c = ID) %>%
  select(id.layer_local_c, all_of(analytes.old.names)) %>%
  rename_with(~analytes.new.names, all_of(analytes.old.names)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "soil fertility data.xlsx") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(tropical.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
tropical.soildata.trans <- transform_values(df = tropical.soildata,
                                       out.name = names(tropical.soildata),
                                       in.name = names(tropical.soildata),
                                       fun.lst = functions.list)

# Final soillab data
tropical.soildata <- tropical.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
tropical.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.qs = path(dir, "ossl_soillab_v2.0.qs")
qs::qsave(tropical.soildata, soillab.qs, preset = "high")
```
Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
tropical.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### Vis-NIR spectra

```{r vis-nir, include=TRUE, echo=TRUE, eval=TRUE}
tropical.visnir <- read_excel(path(dir, "VNIR data.xlsx"))

# Renaming and ID Preparation
tropical.visnir.proc <- tropical.visnir %>%
  as_tibble() %>%
  rename(id.layer_local_c = ID) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# Extracting wavelengths from column names (e.g., "SPC.350" -> 350)
spec.cols <- setdiff(names(tropical.visnir.proc), c("id.layer_local_c", "Field"))
old.wavelengths <- as.numeric(spec.cols)

new.wavelengths <- seq(440, 2150, by = 2)

# Resampling spectra
tropical.visnir.resampled <- tropical.visnir.proc %>%
  select(id.layer_local_c, all_of(spec.cols)) %>%
  mutate(across(all_of(spec.cols), ~ as.numeric(.x) / 100)) %>% 
  { 
    temp_data <- .
    spectra_matrix <- as.matrix(select(temp_data, all_of(spec.cols)))
    resampled_mat <- prospectr::resample(
      X = spectra_matrix, 
      wav = old.wavelengths, 
      new.wav = new.wavelengths, 
      interpol = "spline"
    )
    as_tibble(resampled_mat) %>%
      bind_cols(temp_data %>% select(id.layer_local_c), .)
  } %>%
  select(id.layer_local_c, as.character(new.wavelengths))

# Gaps Analysis
scans.na.gaps <- tropical.visnir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols(tropical.visnir.resampled %>% select(id.layer_local_c), .)

# Extreme negative checks
scans.extreme.neg <- tropical.visnir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols(tropical.visnir.resampled %>% select(id.layer_local_c), .)

# Extreme positive checks
scans.extreme.pos <- tropical.visnir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 1.5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols(tropical.visnir.resampled %>% select(id.layer_local_c), .)

# Consistency summary
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

# Display problematic scans count
scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Final column renaming for OSSL standard
final.visnir.names <- paste0("scan_visnir.", new.wavelengths, "_ref")
tropical.visnir.proc_final <- tropical.visnir.resampled %>%
  rename_with(~final.visnir.names, as.character(new.wavelengths))

# Preparing metadata
tropical.visnir.metadata <- tropical.visnir.proc_final %>%
  select(id.layer_local_c) %>%
  mutate(id.scan_local_c = id.layer_local_c,
         scan.visnir.model.name_utf8_txt = "Veris MSP3 (Veris Technologies, Salina, Kansas, USA)", 
         scan.visnir.license.title_ascii_txt = "CC-BY 4.0",
         scan.visnir.method.optics_any_txt = "",
         scan.visnir.method.preparation_any_txt = "loose powder soil samples (air-dry and grain size ≤ 2 mm).",
         scan.visnir.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/",
         scan.visnir.doi_idf_url = "https://data.mendeley.com/datasets/88c5kvmgbf/1",
         scan.visnir.contact.name_utf8_txt = "Tiago Rodrigues Tavares",
         scan.visnir.contact.email_ietf_txt = "tiagosrt@usp.br")

# Final preparation
tropical.visnir.export <- tropical.visnir.metadata %>%
  left_join(tropical.visnir.proc_final, by = "id.layer_local_c") %>%
  mutate(across(starts_with("id."), as.character))

# Saving
soilvisnir.qs = path(dir, "ossl_visnir_v2.0.qs")
qs::qsave(tropical.visnir.export, soilvisnir.qs, preset = "high")

```


### Quality control for Vis-NIR

The final table must be joined as follows:

- Vis-NIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_visnir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
tropical.availability <- tropical.visnir.export %>%
  select(id.layer_local_c, scan_visnir.440_ref) %>%
  left_join({tropical.metadata %>%
      rename(id.layer_local_c = ID) %>%
      mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
      select(id.layer_local_c, Clay, CEC, pH)}, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information summary
# This tells us how many samples have spectra vs. lab/site data
tropical.availability %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Checking for duplicate SAMPLE_TDR_IDs
tropical.availability %>%
  mutate(across(everything(), as.character)) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n(), .groups = 'drop') %>%
  group_by(column, repeats) %>%
  summarise(count = n(), .groups = 'drop')

```

Soil analytical data summary for Vis-NIR. Note: many scans could not be linked with the wetchem.
```{r summary_visnir, include=TRUE, echo=TRUE, eval=TRUE}
tropical.soildata %>%
  filter(id.layer_local_c %in% tropical.visnir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```
Vis-NIR spectral visualization (100 random spectra):
```{r visnir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
tropical.visnir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_visnir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavelength", values_to = "reflectance") %>%
  dplyr::mutate(wavelength = gsub("scan_visnir.|_ref", "", wavelength)) %>%
  dplyr::mutate(wavelength = as.numeric(wavelength)) %>%
  ggplot(aes(x = wavelength, y = reflectance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(350, 2500, by = 250))+
  labs(title = "Vis-NIR Spectra (100 random scans)",
       x = "Wavelength (nm)",
       y = "Reflectance")+
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```

## References
