---
title: "Serbian dataset preparation for the OSSL"
author: 
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Ran Zhi
    email: jsafanelli@woodwellclimate.org
  - name: Tomislav Hengl
    email: tom.hengl@opengeohub.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for standardizing and importing the Serbian Soil Spectral Library.

Website: [Soil Spectroscopy for Global Good](https://soilspectroscopy.org)  \
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## Original data

Site data, soil lab data, and Mid-Infrared Spectra (MIR) from the Serbia.

It is composed of 135 samples shared by University of Novi Sad and published in @Jovi2019. The soil samples were sampled in 2009. Samples were taken from three different types of soil: arable, meadow, and forest. 3 different depths: 0-30 cm, 30-60 cm and 60-90 cm in each location (n=45).

Wet chemistry analyses include texture, carbon total, carbon organic, carbon inorganic, pH, C/N ratio, and CaCO3. Only texture has complete data for all depths, the other soil properties were determined only for the 0-30 cm layer. The spectra in the MIR range were taken for all depths using a ThermoFisher Nicolet Nexus 670 with a DTGS detector, diffuse reflectance module, and gold as background reference. MIR spectra were obtained in the spectral range of 400-4000 cm-1 in duplicates. The coordinates of the sampling sites are given with lower precision.

Input files:  
- `SSSL.xlsx`: excel file with soil/site, MIR, and ViSNIR (1100-2500 nm, not imported).

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "data.table", "lubridate", "units", "readxl", "stringr",
              "googledrive", "googlesheets4", "tmap", "sf", "rnaturalearth",
              "skimr", "prospectr", "tictoc", "nanoparquet", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
gs4_auth(path = "~/secrets/soilcarbon-soilspec-845d0d3f2fbe.json")
options(scipen = 999)
```

Directory/folder path with original files (not uploaded to GitHub). 
```{r dir}
# dir = "./"
dir = "~/mnt-ossl-private/database/datasets/SRB/"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r soilsite}
serbia.info <- read_xlsx(paste0(dir, "/SSSL.xlsx"), sheet = 1)

# Formatting to OSSL standard
serbia.sitedata <- serbia.info %>%
  select(sample_id, location_use, longitude, latitude, soil_type, depth_cm) %>%
  mutate(location_use = gsub(" - |- | -", "-", location_use)) %>%
  separate(depth_cm, into = c("layer.upper.depth_usda_cm",
                              "layer.lower.depth_usda_cm"), sep = "-") %>%
  mutate(layer.upper.depth_usda_cm = as.numeric(layer.upper.depth_usda_cm),
         layer.lower.depth_usda_cm = as.numeric(layer.lower.depth_usda_cm)) %>%
  rename(id.layer_local_c = sample_id,
         site.location.use_src_txt = location_use,
         longitude.point_wgs84_dd = longitude,
         latitude.point_wgs84_dd = latitude,
         pedon.taxa_src_txt = soil_type) %>%
  # mutate(observation.ogc.schema.title_ogc_txt = 'Open Soil Spectroscopy Library',
  #        observation.ogc.schema_idn_url = 'https://soilspectroscopy.github.io',
  #        observation.date.begin_iso.8601_yyyy.mm.dd = "2019-01-01",
  #        observation.date.end_iso.8601_yyyy.mm.dd = "2019-12-31",
  #        layer.texture_usda_txt = "",
  #        longitude.county_wgs84_dd = NA,
  #        latitude.county_wgs84_dd = NA,
  #        location.country_iso.3166_txt = "SRB",
  #        location.point.error_any_m = 111,
  #        surveyor.title_utf8_txt = "Branislav Jović",
  #        surveyor.contact_ietf_email = "branislav.jovic@dh.uns.ac.rs",
  #        surveyor.address_utf8_txt = 'Faculty of Science, University of Novi Sad, Serbia',
  #        dataset.title_utf8_txt = 'Jović et al. (2019)',
  #        dataset.owner_utf8_txt = 'Jović et al. (2019)',
  #        dataset.code_ascii_txt = 'SERBIA.SSL',
  #        dataset.address_idn_url = 'https://doi.org/10.1016/j.saa.2018.08.039',
  #        dataset.license.title_ascii_txt = 'CC-BY',
  #        dataset.license.address_idn_url = 'https://creativecommons.org/licenses/by/4.0/legalcode',
  #        dataset.doi_idf_url = 'https://doi.org/10.1016/j.saa.2018.08.039',
  #        dataset.contact.name_utf8_txt = "Branislav Jović",
  #        dataset.contact_ietf_email = "branislav.jovic@dh.uns.ac.rs") %>%
  mutate(dataset.code_ascii_txt = "SRB",
         id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .before = 1) %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
site.exp.file = path(dir, "ossl_soilsite_v1.3")
readr::write_csv(serbia.sitedata, str_c(site.exp.file, ".csv.gz"))
nanoparquet::write_parquet(serbia.sitedata, str_c(site.exp.file, ".parquet"))
```

Plotting map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

ocean <- ne_download(scale = 110, type = "ocean", category = "physical", returnclass = "sf")

points <- serbia.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")
tm_shape(ocean) +
  tm_polygons(fill = "lightblue", col = NA) +
  tm_shape(World) +
  tm_polygons(fill = "#f0f0f0", fill_alpha = 0.5, col_alpha = 0.5) +
  tm_shape(points) +
  tm_dots(fill = "firebrick") +
  tm_crs("ESRI:54030") +
  tm_layout(frame = FALSE)
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template for scripted column standardization. Just run once for getting the original names of soil properties, descriptions, data types, and units. Then upload to Google Sheet for editing and manually defining the rules for integrating with the OSSL. Requires Google authentication. A copy of the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwritting, then rename online to download locally.**

```{r soil_template, echo=TRUE, eval=FALSE}
# Checking shared files
list.files(dir)

# Checking column description
serbia.soil <- read_xlsx(paste0(dir, "/SSSL.xlsx"), sheet = 1)

soillab.names <- serbia.soil %>%
  names(.) %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'SSSL.xlsx', .before = 1) %>%
  dplyr::mutate(import = '', ossl_name = '', .after = original_name) %>%
  dplyr::mutate(comment = '')

readr::write_csv(soillab.names, paste0(getwd(), "/serbia_soillab_names.csv"))

# Uploading to google sheet

# FACT CIN folder. Get ID for soildata importing table
googledrive::drive_ls(as_id("0AHDIWmLAj40_Uk9PVA"))

OSSL.soildata.importing <- "19LeILz9AEnKVK7GK0ZbK3CCr2RfeP-gSWn5VpY8ETVM"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names
upload <- dplyr::as_tibble(soillab.names)

# Uploading
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the edited standardization template is saved to this dataset folder.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "SRB")

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:
```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F) %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "ossl_")))

knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:
```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
# Reading soil information
serbia.info <- read_xlsx(path(dir, "/SSSL.xlsx"), sheet = 1)

# Harmonization of names and units
analytes.old.names <- transvalues %>%
  pull(original_name)

analytes.new.names <- transvalues %>%
  pull(ossl_name)

# Selecting and renaming
serbia.soildata <- serbia.info %>%
  rename(id.layer_local_c = sample_id) %>%
  select(id.layer_local_c, all_of(analytes.old.names)) %>%
  rename_with(~analytes.new.names, all_of(analytes.old.names)) %>%
  mutate_at(vars(-id.layer_local_c), as.numeric) %>%
  as.data.frame()

# Removing duplicates
serbia.soildata %>%
  group_by(id.layer_local_c) %>%
  summarise(repeats = n()) %>%
  group_by(repeats) %>%
  summarise(count = n())

# Getting the formulas
functions.list <- transvalues %>%
  mutate(ossl_name = factor(ossl_name, levels = names(serbia.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
serbia.soildata.trans <- transform_values(df = serbia.soildata,
                                          out.name = names(serbia.soildata),
                                          in.name = names(serbia.soildata),
                                          fun.lst = functions.list) %>%
  mutate(oc_usda.c729_w.pct = c.tot_usda.a622_w.pct-0.12*caco3_usda.a54_w.pct)

# Final soillab data
serbia.soildata <- serbia.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
serbia.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())

# Saving version to dataset root dir
soillab.exp.file = path(dir, "ossl_soillab_v1.3")
readr::write_csv(serbia.soildata, str_c(soillab.exp.file, ".csv.gz"))
nanoparquet::write_parquet(serbia.soildata, str_c(soillab.exp.file, ".parquet"))
```

Soil lab data summary.
```{r soil_lab_summary, include=TRUE, echo=TRUE, eval=TRUE}
serbia.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### Mid-infrared spectra

```{r mir, include=TRUE, echo=TRUE, eval=TRUE}
# Dataset
serbia.spec <- read_xlsx(paste0(dir, "/SSSL.xlsx"), sheet = 2)

# Renaming
serbia.spec <- serbia.spec %>%
  rename(id.layer_local_c = sample_id)

# Need to resample spectra
old.wavenumber <- na.omit(as.numeric(names(serbia.spec)))
new.wavenumbers <- seq(600, 4000, by = 2)

serbia.mir <- serbia.spec %>%
  select(-id.layer_local_c) %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavenumber,
                      new.wav = new.wavenumbers, interpol = "spline") %>%
  as_tibble() %>%
  bind_cols({serbia.spec %>%
      select(id.layer_local_c)}, .) %>%
  select(id.layer_local_c, as.character(new.wavenumbers))

# Gaps
scans.na.gaps <- serbia.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(proportion_NA = .) %>%
  bind_cols({serbia.mir %>% select(id.layer_local_c)}, .)

# Extreme negative - irreversible erratic patterns
scans.extreme.neg <- serbia.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x < -1, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_lower0 = .) %>%
  bind_cols({serbia.mir %>% select(id.layer_local_c)}, .)

# Extreme positive, irreversible erratic patterns
scans.extreme.pos <- serbia.mir %>%
  select(-id.layer_local_c) %>%
  apply(., 1, function(x) {round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)}) %>%
  tibble(proportion_higherAbs5 = .) %>%
  bind_cols({serbia.mir %>% select(id.layer_local_c)}, .)

# Consistency summary - problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming
old.wavenumbers <- seq(600, 4000, by = 2)
new.wavenumbers <- paste0("scan_mir.", old.wavenumbers, "_abs")

serbia.mir <- serbia.mir %>%
  rename_with(~new.wavenumbers, as.character(old.wavenumbers)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c))

# # Preparing metadata
# serbia.mir.metadata <- serbia.mir %>%
#   select(id.layer_local_c) %>%
#   mutate(id.scan_local_c = id.layer_local_c) %>%
#   mutate(scan.mir.date.begin_iso.8601_yyyy.mm.dd = ymd("2019-01-01"),
#          scan.mir.date.end_iso.8601_yyyy.mm.dd = ymd("2019-12-31"),
#          scan.mir.model.name_utf8_txt = "ThermoFisher Nicolet Nexus 670",
#          scan.mir.model.code_any_txt = "ThermoFisher_Nicolet_Nexus_670",
#          scan.mir.method.optics_any_txt = "DTGS detector; DRIFT accessory; Gold background",
#          scan.mir.method.preparation_any_txt = "Finelly ground <80 mesh",
#          scan.mir.license.title_ascii_txt = "CC-BY",
#          scan.mir.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/",
#          scan.mir.doi_idf_url = 'https://doi.org/10.1016/j.saa.2018.08.039',
#          scan.mir.contact.name_utf8_txt = "Branislav Jović",
#          scan.mir.contact.email_ietf_txt = "branislav.jovic@dh.uns.ac.rs")
# 
# # Final preparation
# serbia.mir.export <- serbia.mir.metadata %>%
#   left_join(serbia.mir, by = "id.layer_local_c") %>%
#   mutate_at(vars(starts_with("id.")), as.character)

# Saving version to dataset root dir
mir.exp.file = path(dir, "ossl_mir_v1.3")
readr::write_csv(serbia.mir, str_c(mir.exp.file, ".csv.gz"))
nanoparquet::write_parquet(serbia.mir, str_c(mir.exp.file, ".parquet"))
```

## Quality control

The final table must be joined as follows:

- MIR is used as first reference for pairing with soil data.
- Soil lab data are left joined to MIR. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
serbia.availability <- serbia.mir %>%
  select(id.layer_local_c, scan_mir.600_abs) %>%
  left_join(serbia.soildata, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information from serbia
serbia.availability %>%
  mutate_all(as.character) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check - Duplicates are dropped
serbia.availability %>%
  mutate_all(as.character) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n()) %>%
  group_by(column, repeats) %>%
  summarise(count = n())
```

MIR spectral visualization (100 random spectra):
```{r mir_plot, include=TRUE, echo=TRUE, eval=TRUE}
serbia.mir %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_mir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavenumber", values_to = "absorbance") %>%
  dplyr::mutate(wavenumber = gsub("scan_mir.|_abs", "", wavenumber)) %>%
  dplyr::mutate(wavenumber = as.numeric(wavenumber)) %>%
  ggplot(aes(x = wavenumber, y = absorbance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = c(600, 1200, 1800, 2400, 3000, 3600, 4000),
                     transform = "reverse") +
  labs(x = bquote("Wavenumber"~(cm^-1)), y = "Absorbance") +
  theme_light()
```

```{r end}
toc()
rm(list = ls())
gc()
```

## References
