---
title: "Austrian NIR soil spectral library (AUT) for soil health assessments"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 3
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Austrian NIR soil spectral library for soil health assessments (AUT) into the Open Soil Spectral Library.

Website: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## Original data

Site data, Soil lab data, and Near-Infrared (NIR) data from all environmental zones of Austria. 

Further information of the dataset can be found in detail at @fohrafellner_austrian_2026.

Original files:\
- `Austrian NIR Soil Spectral Library_V4.xlsx`: xlsx file with site information, soil information, and NIR spectral data.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "data.table", "lubridate", "units", "readxl", "stringr",
              "googledrive", "googlesheets4", "tmap", "sf", "rnaturalearth",
              "skimr", "prospectr", "tictoc", "nanoparquet", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
gs4_auth(path = "~/secrets/soilcarbon-soilspec-845d0d3f2fbe.json")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r dir_def}
# dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/Austrian"
dir = "~/mnt-ossl-private/database/datasets/AUT"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r soil_site}
austrian.metadata <- read_excel(path(dir, "Austrian NIR Soil Spectral Library_V4.xlsx"),
                            sheet = "Dataset")

austrian.sitedata <- austrian.metadata %>%
  select(Sample_number, Sampling_year, Experiment_number, Sample_source,
         Municipality_code, Environmental_zone, Land_use_type, 
         Sampling_depth_from, Sampling_depth_to) %>%
  rename(id.layer_local_c = Sample_number,
         observation.date_src_yyyy = Sampling_year,
         observation.source_src_txt = Sample_source,
         loc.municipality_src_code = Municipality_code,
         loc.env.zone_src_txt = Environmental_zone,
         loc.experiment_src_txt = Experiment_number,
         site.land.use_src_txt = Land_use_type,
         layer.upper.depth_usda_cm = Sampling_depth_from,
         layer.bottom.depth_usda_cm = Sampling_depth_to) %>%
  mutate(layer.sequence_usda_uint16 = ifelse(layer.upper.depth_usda_cm == 0, 1, 2)) %>%
  # mutate(id.project_ascii_txt = "Austrian NIR Soil Spectral Library for Soil Health Assessments",
  #        observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
  #        observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
  #        dataset.title_utf8_txt = "Austrian NIR Soil Spectral Library for Soil Health Assessments",
  #        dataset.owner_utf8_txt = "European Joint Program for SOIL 'Towards climate-smart sustainable management of agricultural soils' (EJP SOIL) funded by the European Union Horizon 2020 research and innovation programme (Grant Agreement N° 862695)",
  #        dataset.doi_idf_url = "https://zenodo.org/records/17941270",
  #        dataset.license.title_ascii_txt = "CC-BY",
  #        dataset.license.address_idn_url = "https://creativecommons.org/licenses/by/4.0/legalcode",
  #        dataset.contact.name_utf8_txt = "Julia, Fohrafellner",
  #        dataset.contact_ietf_email = "julia.fohrafellner@ages.at") %>%
  mutate(dataset.code_ascii_txt = "AUT",
         id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .before = 1) %>%
  mutate(across(starts_with("id."), as.character))

# Saving version to dataset root dir
site.exp.file = path(dir, "ossl_soilsite_v1.3")
readr::write_csv(austrian.sitedata, str_c(site.exp.file, ".csv.gz"))
nanoparquet::write_parquet(austrian.sitedata, str_c(site.exp.file, ".parquet"))
```

Plotting map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

ocean <- ne_download(scale = 110, type = "ocean", category = "physical", returnclass = "sf")

austria <- World[World$name == "Austria", ]

tmap_mode("plot")
tm_shape(ocean) +
  tm_polygons(fill = "lightblue", col = NA) +
  tm_shape(World) +
  tm_polygons(fill = "#f0f0f0", fill_alpha = 0.5, col_alpha = 0.5) +
  tm_shape(austria) +
  tm_polygons(fill = "firebrick") +
  tm_crs("ESRI:54030") +
  tm_layout(frame = FALSE)
```

### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.

**Always leave the sheet name as TEMP to avoid overwritting, then rename online to dataset code and download a copy for this repository.**

```{r soil_lab_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- austrian.metadata %>%
  select(Sample_number, SOC, SOC_to_clay_ratio,
         TC, Labile_carbon, CaCO3, TN, Phosphorus, pH_CaCl2, pH_Acetate, 
         CEC, Sand, Silt, Clay) %>%
  rename(id.layer_local_c = Sample_number) %>%
  names() %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'Austrian NIR Soil Spectral Library_V4.xlsx', .before = 1) %>%
  dplyr::mutate(import = '', original_unit = '', original_method = '',
                comment = '', ossl_abbrev = '', ossl_method = '',
                ossl_unit = '', ossl_convert = '', ossl_name = '',
                .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the edited standardization template is saved to this dataset folder.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "AUT")

# Saving to folder
readr::write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)

 transvalues <- transvalues%>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))
 
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
austrian.reference <- austrian.metadata

# Harmonization of names and units
valid_mappings <- transvalues %>%
  filter(table == "Austrian NIR Soil Spectral Library_V4.xlsx") %>%
  filter(!is.na(ossl_name) & ossl_name != "") 

analytes.old.names <- valid_mappings %>% pull(original_name)
analytes.new.names <- valid_mappings %>% pull(ossl_name)

#analytes.old.names.clean <- analytes.old.names[analytes.old.names != "Sample_number"]
#analytes.new.names.clean <- analytes.new.names[analytes.old.names != "Sample_number"]

austrian.soildata <- austrian.reference %>%
  rename(id.layer_local_c = Sample_number) %>%
  select(id.layer_local_c, all_of(analytes.old.names)) %>%
  rename_with(~analytes.new.names, all_of(analytes.old.names)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- transvalues %>%
  filter(table == "Austrian NIR Soil Spectral Library_V4.xlsx") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(austrian.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
austrian.soildata.trans <- transform_values(df = austrian.soildata,
                                       out.name = names(austrian.soildata),
                                       in.name = names(austrian.soildata),
                                       fun.lst = functions.list)

# Final soillab data
austrian.soildata <- austrian.soildata.trans %>%
  mutate_at(vars(starts_with("id.")), as.character)

# Checking total number of observations
austrian.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.exp.file = path(dir, "ossl_soillab_v1.3")
readr::write_csv(austrian.soildata, str_c(soillab.exp.file, ".csv.gz"))
nanoparquet::write_parquet(austrian.soildata, str_c(soillab.exp.file, ".parquet"))
```

Soil lab data summary.
```{r soil_lab_summary, include=TRUE, echo=TRUE, eval=TRUE}
austrian.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### NIR spectra

Authors describe in the paper that the spectra was measured in reflectance units, then transformed to pseudo absorbance. We will convert back to reflectance following OSSL specifications for VisNIR and NIR.

```{r nir, include=TRUE, echo=TRUE, eval=TRUE}
# Renaming
austrian.nir.proc <- austrian.metadata %>%
  rename(id.layer_local_c = Sample_number) %>%
  select(id.layer_local_c, `680`:last_col())

# Need to resample spectra
spec.cols <- names(austrian.nir.proc)[-1] # Remove the ID column
old.wavelengths <- as.numeric(spec.cols)
new.wavelengths <- seq(680, max(old.wavelengths), by = 2)

# Resampling spectra
austrian.nir.resampled <- austrian.nir.proc %>%
  column_to_rownames("id.layer_local_c") %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavelengths,
                      new.wav = new.wavelengths, interpol = "spline") %>%
  as_tibble(rownames = "id.layer_local_c")

# Backtransforming to reflectance
austrian.nir.resampled <- austrian.nir.resampled %>%
  mutate(across(all_of(as.character(new.wavelengths)), ~1/10^.x))

# Check for NAs
scans.na.gaps <- austrian.nir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(id.layer_local_c = austrian.nir.resampled$id.layer_local_c, proportion_NA = .)

# Check for extreme negative values 
scans.extreme.neg <- austrian.nir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(1, function(x) round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)) %>%
  tibble(id.layer_local_c = austrian.nir.resampled$id.layer_local_c, proportion_lower0 = .)

# Check for extreme positive values 
scans.extreme.pos <- austrian.nir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(1, function(x) round(100*(sum(x > 1, na.rm=TRUE))/(length(x)), 2)) %>%
  tibble(id.layer_local_c = austrian.nir.resampled$id.layer_local_c, proportion_higherRef1 = .)

# Summary of problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

scans.summary %>%
  select(-id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "check", values_to = "value") %>%
  filter(value > 0) %>%
  group_by(check) %>%
  summarise(count = n())

# Renaming and Metadata
final.visnir.names <- paste0("scan_nir.", new.wavelengths, "_ref")

austrian.nir.final <- austrian.nir.resampled %>%
  rename_with(~final.visnir.names, as.character(new.wavelengths))

# # Customizing metadata for the Austrian dataset
# austrian.nir.metadata <- austrian.nir.final %>%
#   select(id.layer_local_c) %>%
#   mutate(
#     id.scan_local_c = id.layer_local_c,
#     scan.nir.date.begin_iso.8601_yyyy = ymd("1998-01-01"), 
#     scan.visnir.date.end_iso.8601_yyyy = ymd("2023-12-31"), 
#     scan.nir.model.name_utf8_txt = "SpectraStarTM XL near-infrared spectrometer from Unity Scientific (Brookfield, CT, USA)", 
#     scan.nir.license.title_ascii_txt = "CC-BY",
#     scan.nir.contact.name_utf8_txt = "Austrian Dataset Lead",
#     scan.nir.doi_idf_url = "https://zenodo.org/records/17941270",
#     scan.nir.contact.name_utf8_txt = "Julia, Fohrafellner",
#     scan.nir.contact.email_ietf_txt = "julia.fohrafellner@ages.at"
#   )

# Export

# austrian.nir.export <- austrian.nir.metadata %>%
#   left_join(austrian.nir.final, by = "id.layer_local_c")

# Save as .qs file
nir.exp.file = path(dir, "ossl_nir_v1.3")
readr::write_csv(austrian.nir.final, str_c(nir.exp.file, ".csv.gz"))
nanoparquet::write_parquet(austrian.nir.final, str_c(nir.exp.file, ".parquet"))
```

### Quality control for NIR

The final table must be joined as follows:

- NIR is used as first reference for pairing with site and soil data.
- Site and soil lab data are left joined to nir. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_nir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
austrian.availability <- austrian.nir.final %>%
  select(id.layer_local_c, scan_nir.680_ref) %>%
  left_join(austrian.soildata, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information summary
# This tells us how many samples have spectra vs. lab/site data
austrian.availability %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check 
austrian.availability %>%
  mutate(across(everything(), as.character)) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n(), .groups = 'drop') %>%
  group_by(column, repeats) %>%
  summarise(count = n(), .groups = 'drop')
```

Soil analytical data summary with NIR. Note: many scans could not be linked with some of the wetchem.
```{r summary_nir, include=TRUE, echo=TRUE, eval=TRUE}
austrian.availability %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

NIR spectral visualization (100 random spectra).\

**PLEASE NOTE the bad interpolation between the detectors switch around 1350 nm. Unfortunately, simple splice correction does not work in this case.**

```{r nir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
austrian.nir.final %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_nir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavelength", values_to = "reflectance") %>%
  dplyr::mutate(wavelength = gsub("scan_nir.|_ref", "", wavelength)) %>%
  dplyr::mutate(wavelength = as.numeric(wavelength)) %>%
  ggplot(aes(x = wavelength, y = reflectance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(680, 2500, by = 200))+
  labs(title = "NIR spectra (100 random scans)",
       x = "Wavelength (nm)",
       y = "Reflectance")+
  theme_light()
```

```{r}
toc()
rm(list = ls())
gc()
```

## References


