---
title: "Importing VNIR KSSL data"
author: Tomislav Hengl (tom.hengl@opengeohub.org), Wanderson Mendes de Sousa (wanderson.mendes@opengeohub.org)
  and Leandro Parente (leandro.parente@opengeohub.org)
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Kellogg Soil Survey Laboratory inputs
Part of: [SoilSpec4GG](https://gitlab.com/soilspec4gg)  
Project: [Soil Spectroscopy for Global Good](https://soilspectroscopy.org)  
Last update:  `r Sys.Date()`  
Download files: [diskstation.opengeohub.org](http://gofile.me/6J25n/uIkbNZ5jp)  

```{r, include=FALSE}
options(warn=0)
```

This R Markdown presents quick and short code to import and the visible and near-infrared spectra from KSSL library. Dataset properties are explained 
in detail in [@wijewardane2018predicting] and [@sanderman2020mid].

Input datasets:

- `VNIR_Spectral_Library`: Entire folder containing all asd files with VNIR soil spectral reflectances;
- `KSSL_DB_Access2000.mdb`: original DB as Microsoft Access MDB file;  

For the DB structure and use refer to "Introduction to the KSSL Laboratory Information Management System" contacts: Rich Ferguson & Scarlett Murphy (NRCS USDA).

Loading libraries:
```{r, echo=FALSE}
if(!require(Hmisc)){install.packages("Hmisc"); require(Hmisc)}
if(!require(asdreader)){install.packages("asdreader"); require(asdreader)}
if(!require(plyr)){install.packages("plyr"); require(plyr)}
if(!require(dplyr)){install.packages("dplyr"); require(dplyr)}

```

#### Setting the folder path first of all

```{r}
dir = "/mnt/europa/SoilSpec/SoilSpec4GG/R_Codes/KSSL/"
```

#### _1. Importing mdb access file:_
```{r}
## Importing the access file/ see imported file into KSSL_README is the same as `kssl.df`
kssl.df = mdb.get(paste(dir, "KSSL_DB_Access2000.mdb", sep = ""))
names(kssl.df)

```

#### _2. Getting VNIR IDs and fields from `kssl.df` and then joining all VNIR info:_
```{r}
vnir.mas = kssl.df$vnir_scan_mas_data
colnames(vnir.mas)
vnir.det = kssl.df$vnir_scan_det_data
colnames(vnir.det)
vnir = plyr::join(vnir.det, vnir.mas)
head(vnir, 5)
```

#### _3. Importing, reading and combining asd files into one dataframe:_

Setting the folder location and create a list with the files.
```{r}
my_dir = paste(dir, "VNIR_Spectral_Library/", sep = "")

## Grab the list of files from different subfolders
vnir.data.files = list.files(path = my_dir, 
                             recursive = T,
                             pattern = ".asd",
                             full.names = T)
head(vnir.data.files, 5)
```

Reading ASD files through a function using `get_spectra` from `asdreader` R package:

```{r}
vnir.read.asd.files = function(x){
  dat.read.asd = get_spectra(x)
}

## Taking the list of files as input, passing through the function and returning output in vector or matrix. 
vnir.df.raw = sapply(vnir.data.files, vnir.read.asd.files)
class(vnir.df.raw)
```

Converting matrix to dataframe and transposing it:

```{r}
vnir.df.raw = as.data.frame(t(vnir.df.raw))
head(vnir.df.raw[, c(1:2)], 5)
```
Check single file to get bands' names
Get the path to the demo file
```{r}
asd.tf <- paste(my_dir, "C2001USCA066/27038MD01.asd", sep = "")
soil.spec <- get_spectra(asd.tf)
```

Example with single file name:
```{r}
matplot(t(soil.spec), type = 'l', xlab = "Wavelength (nm)", ylab = "Reflectance factor (x100)")
```

Getting the wavelength numbers and changing colnames in the `vnir.df.raw` table which contains the raw VNIR data:

```{r}
vnir.wav = colnames(soil.spec)
colnames(vnir.df.raw) = vnir.wav
```

Getting the id from path folder to link with the data from .mdb tables.

```{r}
colnames(vnir)
head(vnir$scan.path.name, 5)
```

```{r}
vnir.df.raw$scan.path.name = rownames(vnir.df.raw)
vnir.df.raw$scan.path.name = sub(".*/", "", vnir.df.raw$scan.path.name)
vnir.df.raw = vnir.df.raw %>%
  select("scan.path.name", everything())
head(vnir.df.raw$scan.path.name, 5)
```

```{r}
#names(vnir.df.raw) <- gsub(x = names(vnir.df.raw), pattern = "V", replacement = "")
vnir.ok = plyr::join(vnir, vnir.df.raw)
colnames(vnir.ok)[2162]### Example with single file name
```

Standarising the VNIR colnames:

```{r}
colnames(vnir.ok)[12:2162] <- paste("spec.vnir", colnames(vnir.ok)[12:2162], sep = "_")
head(vnir.ok[, c(1:5, 12:14)], 5)
```

#### _Exporting only spectra_

```{r}
#colnames(vnir.ok)
#saveRDS(vnir.ok, paste(dir, "vnir_09MAR2021.rds", sep = ""))
```

## Quality control

Removing duplicates:
```{r}
drop.col = c("vnir.scan.det.id", "rep.num", "light.source", "scan.date", 
             "scan.path.name", "vnir.scan.mas.id", "anal.batch.id", 
             "batch.ord.pos", "caco3.predict", "caco3.mdistance")

vnir.ok = vnir.ok[, !(names(vnir.ok) %in% drop.col)] #raw data 69715 obs

vnir.ok <- vnir.ok %>%
  group_by(smp.id) %>%
  summarise_all(mean)

head(vnir.ok[, 1:10], 5)
```

Linking the `lay.id` wih `smp.id` from mdb access sample table
```{r}
vnir.ok$lay_id = plyr::join(vnir.ok[,"smp.id"], kssl.df$sample, by="smp.id", match = "first")$lay.id

vnir.ok = vnir.ok %>%
  select(c("lay_id", "smp.id"), everything())


## Linking the unique ID from `kssl.yw`
kssl.yw = readRDS(paste(dir, "KSSL_soillab.rds", sep = ""))
vnir.refl = plyr::join(vnir.ok, kssl.yw[, c("unique_id", "lay_id")])

vnir.refl = vnir.refl %>%
  select(c("unique_id", "lay_id", "smp.id"), everything())

vnir.refl.ok = vnir.refl[!is.na(vnir.refl$unique_id), ] #31054 observations with soil data

```

Removing spectral response under 100%:
(note: the reflectance ranges from 0 to 100%.)

```{r}
vnir.refl.plot = vnir.refl.ok[vnir.refl.ok$spec.vnir_350 < 1,]
vnir.refl.plot = vnir.refl.ok[vnir.refl.ok$spec.vnir_1500 < 1,]
vnir.refl.plot = vnir.refl.ok[vnir.refl.ok$spec.vnir_2500 < 1,]
```

Selecting the soil particles that will be linked to VNIR spectal data:

```{r}
vnir.refl.plot = plyr::join(vnir.refl.plot, kssl.yw[, c("unique_id", "lay_id", "clay.tot_3a1a1a_pctw",
                                                        "sand.tot_3a1a1a_pctw", "silt.tot_3a1a1a_pctw")])

vnir.refl.plot = vnir.refl.plot %>%
  select(c("unique_id", "lay_id", "smp.id", "clay.tot_3a1a1a_pctw", "silt.tot_3a1a1a_pctw", "sand.tot_3a1a1a_pctw"), everything())

row.has.na <- apply(vnir.refl.plot, 1, function(x){any(is.na(x))})
sum(row.has.na) #  13885 observations with NAs
vnir.refl.plot <- vnir.refl.plot[complete.cases(vnir.refl.plot),]

# Selecting low and high values of clay content to grab the Unique Id 
low.v = subset(vnir.refl.plot, clay.tot_3a1a1a_pctw <= 10)
low.v[1:5, 1:5]

high.v = subset(vnir.refl.plot, clay.tot_3a1a1a_pctw >= 60)
high.v[1:20, 1:5]

ids = c("79bc428e-8248-11eb-91f0-0242ac110002", "79be51c8-8248-11eb-91f0-0242ac110002", "79c428be-8248-11eb-91f0-0242ac110002", "79bc27e0-8248-11eb-91f0-0242ac110002")

result = vnir.refl.plot[vnir.refl.plot$unique_id %in% ids,]
```

Grouping in soil texture - 6 levels:

```{r}
st.levels = c(-Inf, 10, 15, 25, 35, 60, Inf)
st.labels = c("very sandy", "sandy", "sandy loam", "clayey loam", "clayey", "very clayey")
result = result %>% mutate(soil.texture = cut(clay.tot_3a1a1a_pctw, st.levels, labels = st.labels))

result = result %>%
  select(c("unique_id", "lay_id", "smp.id", "clay.tot_3a1a1a_pctw", "soil.texture"), everything())

result = result[order(result$clay.tot_3a1a1a_pctw),]
result$soil.texture
```

Plotting the final spectral curves:
(note: Sandy soils has to present high reflectance values while clayey soils has to display low values.)

```{r}
matplot(vnir.wav,
        t(result[1:4, 8:2158]),
        type = 'l',
        #lty = 1:2,
        lwd = 2,
        col=c("burlywood3", "gold2", "red", "black"),
        #ylim = c(0, 0.1),
        xlim = c(350, 2500),
        xlab =  'Wavelength (nm)', #expression(paste('Wavelength(cm'^'-1',')')),
        ylab = 'Reflectance',
        las=1, font=2)
grid(lwd=1, nx=10, ny=10)
legend(aes(1200, 0.28), legend=round(result$clay.tot_3a1a1a_pctw, 1), 
       title = "Clay",
       col= c("burlywood3", "gold2", "red", "black"),
       #lty = 1:2,
       lwd = 2,
       cex = 1,
       box.lty = 0,
       x.intersp = 0.3,
       seg.len = 1.2,
       #ncol = 6,
       bg = "transparent")

legend(aes(1600, 0.28), legend=result$silt.tot_3a1a1a_pctw, 
       title = "Silt",
       col= c("burlywood3", "gold2", "red", "black"),
       #lty = 1:2,
       lwd = 2,
       cex = 1,
       box.lty = 0,
       x.intersp = 0.3,
       seg.len = 1.2,
       #ncol = 6,
       bg = "transparent")

legend(aes(2000, 0.28), legend=result$sand.tot_3a1a1a_pctw, 
       title = "Sand",
       col= c("burlywood3", "gold2", "red", "black"),
       #lty = 1:2,
       lwd = 2,
       cex = 1,
       box.lty = 0,
       x.intersp = 0.3,
       seg.len = 1.2,
       #ncol = 6,
       bg = "transparent")

```

```{r}
save.image(paste(dir, "KSSL_VNIR_Import.RData", sep = ""))
```
## End.