---
title: "Martinique dataset preparation for the OSSL"
author: 
  - name: Ran Zhi
    email: rzhi@woodwellclimate.org
  - name: Jose L. Safanelli
    email: jsafanelli@woodwellclimate.org
  - name: Jonathan Sanderman
    email: jsanderman@woodwellclimate.org
format: 
  gfm:
    toc: true
    toc-depth: 2
bibliography: reference.bib
csl: ../../tex/apa.csl
cite-method: citeproc
link-citations: true
twitter-card:
  site: "@soilspec"
editor: 
  markdown: 
    wrap: 72
---

Code repository for preparing and importing the Martinique NIR Soil Spectral Dataset (MTQ) into the Open Soil Spectral Library.

Website: [Soil Spectroscopy for Global
Good](https://soilspectroscopy.org)\
Development: <https://github.com/soilspectroscopy>\
Last update: `r Sys.Date()`\
Additional documentation:

## The MTQ original data

Site data, Soil lab data, and Near-Infrared (NIR) data from rural areas of the Martinique island. Further information of the dataset can be found in detail at @barthes_soil_2024.

Original files:\
- `SoilCarbonMartinique_v2.tab`: tab file with site information, soil information, and NIR spectral data.

```{r packages, include=FALSE, echo=FALSE, eval=TRUE}
packages <- c("tidyverse", "prospectr", "units", "readxl", "stringr",
              "olctools", "openssl", "tmap", "sf", "skimr", "lubridate",
              "googledrive", "googlesheets4", "data.table", "doMC", "tictoc",
              "qs", "dplyr", "tidyr", "fs")

invisible(lapply(packages, library, character.only = TRUE))

source("../../R/functions/SSL_functions.R")
```

Directory/folder path with original files (not uploaded to GitHub).

```{r}
dir = "/Users/rzhi/Projects/git/ossl-imports-internal/dataset/Martinique"
tic()
```

## Data standardization to the OSSL format

### Site information

```{r}
martinique.metadata <- read_tsv(path(dir, "SoilCarbonMartinique_v2.tab"), locale = locale(encoding = "latin1"))

# Function to convert "14° 49' N" to decimal degrees
dms_to_dd <- function(x) {
  parts <- str_match(x, "([0-9.]+)[^0-9.]+([0-9.]+)[^0-9.]+([NSEW])")
  deg <- as.numeric(parts[,2])
  min <- as.numeric(parts[,3])
  dir <- parts[,4]
  dd <- deg + (min / 60)
  dd <- ifelse(dir %in% c("W", "S"), -dd, dd)
  return(dd)
}

martinique.sitedata <- martinique.metadata %>%
  select(Sample, Region, Site, Latitude, Longitude, `Depth layer`, `Soil type`, 
         `Land use`, `Land-use duration and previous use`) %>%
  rename(id.layer_local_c = Sample) %>%
  mutate(
    latitude.point_wgs84_dd = dms_to_dd(Latitude),
    longitude.point_wgs84_dd = dms_to_dd(Longitude)
  ) %>%
  mutate(depth_temp = gsub(" cm", "", `Depth layer`)) %>%
  separate(depth_temp, into = c("layer.upper.depth_usda_cm", "layer.lower.depth_usda_cm"), sep = "-") %>%
  mutate(across(c(layer.upper.depth_usda_cm, layer.lower.depth_usda_cm), as.numeric)) %>%
  group_by(Site) %>%
  arrange(layer.upper.depth_usda_cm, .by_group = TRUE) %>%
  mutate(layer.sequence_usda_uint16 = row_number()) %>%
  ungroup() %>%
  mutate(id.project_ascii_txt = "Data on soil organic carbon content and stock in Martinique – relations to near infrared spectra",
         dataset.code_ascii_txt = "Martinique.NIR",
         observation.ogc.schema.title_ogc_txt = "Open Soil Spectroscopy Library",
         observation.ogc.schema_idn_url = "https://soilspectroscopy.github.io",
         dataset.title_utf8_txt = "Data on soil organic carbon content and stock in Martinique – relations to NIR",
         dataset.owner_utf8_txt = "Aurélie Cambou (UMR Eco&Sols - Univ. Montpellier, CIRAD, INRAE, IRD, L'Institut Agro Montpellier - France)",
         dataset.doi_idf_url = "https://dataverse.ird.fr/dataset.xhtml?persistentId=doi:10.23708/C2TV6W",
         dataset.license.title_ascii_txt = "CC-BY 4.0",
         dataset.license.address_idn_url = "https://creativecommons.org/licenses/by-nc/4.0/",
         dataset.contact.name_utf8_txt = "Aurélie Cambou",
         dataset.contact_ietf_email = "aurelie.cambou@ird.fr") %>%
  mutate(id.layer_uuid_txt = openssl::md5(paste0(dataset.code_ascii_txt, id.layer_local_c)),
         .after = id.project_ascii_txt) %>%
  mutate(across(starts_with("id."), as.character))

# Saving version to dataset root dir
site.qs = path(dir, "ossl_soilsite_v2.0.qs")
qs::qsave(martinique.sitedata, site.qs, preset = "high")
```

Plotting sites map:
```{r map, include=TRUE, echo=TRUE, eval=TRUE}
data("World")

points <- martinique.sitedata %>%
  filter(!is.na(longitude.point_wgs84_dd),
         !is.na(latitude.point_wgs84_dd)) %>%
  st_as_sf(coords = c('longitude.point_wgs84_dd', 'latitude.point_wgs84_dd'), crs = 4326)

tmap_mode("plot")

tm_shape(World) +
  tm_polygons('#f0f0f0f0', border.alpha = 0.2) +
  tm_shape(points) +
  tm_dots()
```


### Soil lab information (reference analytical data)

NOTE: The code chunk below must be run just once for getting a template
for scripted column standardization. Just run once for getting the
original names of soil properties, descriptions, data types, and units.
Then upload to Google Sheet for editing and manually defining the rules
for integrating with the OSSL. Requires Google authentication. A copy of
the output file is saved to this folder for archiving purposes.

```{r soil_template, echo=TRUE, eval=FALSE}
# Getting soillab original variables

soillab.names <- martinique.metadata %>%
  select(Sample, `SOCg (g/kg soil < 2 mm)`, 
    `Nt (g/kg soil < 2 mm)`, 
    `Clay 0-0.002 mm (% soil < 2 mm)`,
    `Sand 0.5-2 mm (% soil < 2 mm)`, 
    `Sand 0.2-0.5 mm (% soil < 2 mm)`, 
    `Sand 0.05-0.2 mm (% soil < 2 mm)`,
    `Silt 0.02-0.05 mm (% soil < 2 mm)`, 
    `Silt 0.002-0.02 mm (% soil < 2 mm)`,
    `Bulk density (kg total soil/dm3 total soil)`,
    `Coarse particles > 2 mm (% total soil)`,
    `SOCv (kg/dm3 total soil)`) %>%
  rename(id.layer_local_c = Sample) %>%
  names() %>%
  tibble(original_name = .) %>%
  dplyr::mutate(table = 'SoilCarbonMartinique_v2.tab', .before = 1) %>%
  dplyr::mutate(import = '', original_unit = '', original_method = '', comment = '', ossl_abbrev = '', 
                ossl_method = '', ossl_unit = '', ossl_convert = '', ossl_name = '', .after = original_name)

readr::write_csv(soillab.names, path(getwd(), "soillab_original_names.csv"))

# Uploading to google sheet

# Drive 'Open Soil Spectral Library'
# Folder 'Database v2'
googledrive::drive_ls(as_id("1l9VM4fFks2xBloDE69EFTfPgQeFx5aGw"))

# Use the ID of the file 'OSSL_v2_tab2_soildata_importing'
OSSL.soildata.importing <- "1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es"

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)

# Checking readme
googlesheets4::read_sheet(OSSL.soildata.importing, sheet = 'readme')

# Preparing soillab.names for this dataset
upload <- dplyr::as_tibble(soillab.names)

# Uploading with TEMP name. Check online, move to sequence, and update name
googlesheets4::write_sheet(upload, ss = OSSL.soildata.importing, sheet = "TEMP")

# Checking metadata
googlesheets4::as_sheets_id(OSSL.soildata.importing)
```

NOTE: The code chunk below must be run just once. Run for getting the
column standardization rules after editing online on Google Sheets. A
copy of the output file is saved to this folder for archiving purposes.

```{r soilab_download, echo=TRUE, eval=FALSE}
# Downloading from google sheet

# Checking metadata
googlesheets4::as_sheets_id("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es")

# Preparing soillab.names
transvalues <- googlesheets4::read_sheet("1mWTDJDuMp4oObcCxAy9gofSkrkcSWWf1TtEW2SuC1es",
                                         sheet = "Martinique") %>%
  filter(import == TRUE) %>%
  select(contains(c("table", "id", "original_name", "original_unit" , "ossl_")))

# Saving to folder
write_csv(transvalues, path(getwd(), "soillab_standardized_names.csv"))
```

Reading standardization rules:

```{r soilab_transvalues, include=TRUE, echo=TRUE, eval=TRUE}
transvalues <- read_csv(path(getwd(), "soillab_standardized_names.csv"),
                        show_col_types = F)
knitr::kable(transvalues)
```

Standardizing soil data to the OSSL format:

```{r soilab_preparation, include=TRUE, echo=TRUE, eval=TRUE}
# Harmonization of names and units

valid_mappings <- transvalues %>%
  filter(table == "SoilCarbonMartinique_v2.tab") %>%
  filter(!is.na(ossl_name) & ossl_name != "" & !original_name %in% c("Sand.total.new", "Silt.total.new"))

analytes.old.names <- valid_mappings %>% pull(original_name)
analytes.new.names <- valid_mappings %>% pull(ossl_name)

# Initial selection and renaming to OSSL/Temporary names
martinique.soildata <- martinique.metadata %>%
  rename(id.layer_local_c = Sample) %>%
  select(id.layer_local_c, all_of(analytes.old.names)) %>%
  rename_with(~analytes.new.names, all_of(analytes.old.names)) %>%
  mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
  as.data.frame()

# Getting the formulas
functions.list <- valid_mappings %>%
  filter(table == "SoilCarbonMartinique_v2.tab") %>%
  mutate(ossl_name = factor(ossl_name, levels = names(martinique.soildata))) %>%
  arrange(ossl_name) %>%
  pull(ossl_convert) %>%
  c("x", .)

# Applying transformation rules
martinique.soildata.trans <- transform_values(df = martinique.soildata,
                                       out.name = names(martinique.soildata),
                                       in.name = names(martinique.soildata),
                                       fun.lst = functions.list)

# Accumulate Silt and Sand into Total OSSL Columns ---
martinique.soildata <- martinique.soildata.trans %>%
  mutate(
    # Total Sand (Summing the 3 temporary sand columns)
    sand.tot_iso.11277_w.pct = ifelse(
      is.na(sand.m.temporary_iso.11277_w.pct) | 
      is.na(sand.f.temporary_iso.11277_w.pct) | 
      is.na(san.vf.temporary_iso.11277_w.pct), 
      NA, 
      sand.m.temporary_iso.11277_w.pct + sand.f.temporary_iso.11277_w.pct + san.vf.temporary_iso.11277_w.pct
    ),
    # Total Silt (Summing the 2 temporary silt columns)
    silt.tot_iso.11277_w.pct = ifelse(
      is.na(silt.c.temporary_iso.11277_w.pct) | 
      is.na(silt.f.temporary_iso.11277_w.pct), 
      NA, 
      silt.c.temporary_iso.11277_w.pct + silt.f.temporary_iso.11277_w.pct
    )
  )

# Final soillab data
martinique.soildata <- martinique.soildata %>%
  select(
    id.layer_local_c,
    oc_iso.10694_w.pct,
    n.tot_iso.13878_w.pct,
    clay.tot_iso.11277_w.pct,
    silt.tot_iso.11277_w.pct,
    sand.tot_iso.11277_w.pct,
    bd_iso.11277_g.cm3,
    cf_iso.11464_w.pct
  ) %>%
  mutate_at(vars(starts_with("id.")), as.character)


# Checking total number of observations
martinique.soildata %>%
  distinct(id.layer_local_c) %>%
  summarise(count = n())
  
# Saving version to dataset root dir
soillab.qs = path(dir, "ossl_soillab_v2.0.qs")
qs::qsave(martinique.soildata, soillab.qs, preset = "high")
```

Soil lab data summary.
```{r summary, include=TRUE, echo=TRUE, eval=TRUE}
martinique.soildata %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

### NIR spectra
```{r nir, include=TRUE, echo=TRUE, eval=TRUE}

# Renaming
martinique.nir.proc <- martinique.metadata %>%
  rename(id.layer_local_c = Sample) %>%
  select(id.layer_local_c, starts_with("Abs"))

# Filter out samples without spectra
martinique.nir.proc <- martinique.nir.proc %>%
  filter(!is.na(`Abs 1100 nm`))

# Need to resample spectra
spec.cols <- names(martinique.nir.proc)[-1] # Remove the ID column

old.wavelengths <- as.numeric(gsub("Abs | nm", "", spec.cols))
new.wavelengths <- seq(1100, max(old.wavelengths), by = 2)

# Resampling spectra
martinique.nir.resampled <- martinique.nir.proc %>%
  column_to_rownames("id.layer_local_c") %>%
  as.matrix() %>%
  prospectr::resample(X = ., wav = old.wavelengths, new.wav = new.wavelengths, interpol = "spline") %>%
  as_tibble(rownames = "id.layer_local_c")

# Check for NAs
scans.na.gaps <- martinique.nir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(1, function(x) round(100*(sum(is.na(x)))/(length(x)), 2)) %>%
  tibble(id.layer_local_c =martinique.nir.resampled$id.layer_local_c, proportion_NA = .)

# Check for extreme negative values 
scans.extreme.neg <- martinique.nir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(1, function(x) round(100*(sum(x < 0, na.rm=TRUE))/(length(x)), 2)) %>%
  tibble(id.layer_local_c = martinique.nir.resampled$id.layer_local_c, proportion_lower0 = .)

# Check for extreme positive values 
scans.extreme.pos <- martinique.nir.resampled %>%
  select(-id.layer_local_c) %>%
  apply(1, function(x) round(100*(sum(x > 5, na.rm=TRUE))/(length(x)), 2)) %>%
  tibble(id.layer_local_c = martinique.nir.resampled$id.layer_local_c, proportion_higherAbs5 = .)

# Summary of problematic scans
scans.summary <- scans.na.gaps %>%
  left_join(scans.extreme.neg, by = "id.layer_local_c") %>%
  left_join(scans.extreme.pos, by = "id.layer_local_c")

# 4. Renaming and Metadata
final.visnir.names <- paste0("scan_nir.", new.wavelengths, "_abs")
martinique.nir.final <- martinique.nir.resampled %>%
  rename_with(~final.visnir.names, as.character(new.wavelengths))

# Customizing metadata for the Martinique dataset
martinique.nir.metadata <- martinique.nir.final %>%
  select(id.layer_local_c) %>%
  mutate(
    id.scan_local_c = id.layer_local_c,
    scan.nir.date.begin_iso.8601_yyyy = ymd("2004-01-01"), 
    scan.visnir.date.end_iso.8601_yyyy = ymd("2005-12-31"), 
    scan.nir.model.name_utf8_txt = "Foss NIRSystems 5000 spectrophotometer (Laurel, MD, USA)", 
    scan.nir.license.title_ascii_txt = "CC-BY 4.0",
    scan.nir.doi_idf_url = "https://dataverse.ird.fr/dataset.xhtml?persistentId=doi:10.23708/C2TV6W",
    scan.nir.contact.name_utf8_txt = "Aurélie Cambou",
    scan.nir.contact.email_ietf_txt = "aurelie.cambou@ird.fr"
  )

# 5. Export
martinique.nir.export <- martinique.nir.metadata %>%
  left_join(martinique.nir.final, by = "id.layer_local_c")

# Save as .qs file
qs::qsave(martinique.nir.export, "ossl_nir_v2.0.qs", preset = "high")

```

### Quality control for NIR

The final table must be joined as follows:

- NIR is used as first reference for left join.
- Then it is left joined with the site and soil lab data. This drop data without any available scan.

The availability of data is summarized below:
```{r bind_test_nir, include=TRUE, echo=TRUE, eval=TRUE}
# Taking a few representative columns for checking the consistency of joins
martinique.availability <- martinique.nir.export %>%
  select(id.layer_local_c, scan_nir.1100_abs) %>%
  left_join({martinique.metadata %>%
      rename(id.layer_local_c = Sample) %>%
      mutate(id.layer_local_c = as.character(id.layer_local_c)) %>%
      select(id.layer_local_c, `SOCg (g/kg soil < 2 mm)`, 
             `Nt (g/kg soil < 2 mm)`, 
             `Clay 0-0.002 mm (% soil < 2 mm)`)}, by = "id.layer_local_c") %>%
  filter(!is.na(id.layer_local_c))

# Availability of information summary
# This tells us how many samples have spectra vs. lab/site data
martinique.availability %>%
  mutate(across(everything(), as.character)) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  filter(!is.na(value)) %>%
  group_by(column) %>%
  summarise(count = n())

# Repeats check 
martinique.availability %>%
  mutate(across(everything(), as.character)) %>%
  select(id.layer_local_c) %>%
  pivot_longer(everything(), names_to = "column", values_to = "value") %>%
  group_by(column, value) %>%
  summarise(repeats = n(), .groups = 'drop') %>%
  group_by(column, repeats) %>%
  summarise(count = n(), .groups = 'drop')

```

Soil analytical data summary for NIR. Note: many scans could not be linked with the wetchem.
```{r summary_nir, include=TRUE, echo=TRUE, eval=TRUE}
martinique.soildata %>%
  filter(id.layer_local_c %in% martinique.nir.export$id.layer_local_c) %>%
  mutate(id.layer_local_c = factor(id.layer_local_c)) %>%
  skimr::skim() %>%
  dplyr::select(-numeric.hist, -complete_rate)
```

NIR spectral visualization (100 random spectra):
```{r nir_plot, include=TRUE, echo=TRUE, eval=TRUE}
set.seed(42)
martinique.nir.export %>%
  sample_n(100) %>%
  select(all_of(c("id.layer_local_c")), starts_with("scan_nir.")) %>%
  tidyr::pivot_longer(-all_of(c("id.layer_local_c")),
                      names_to = "wavelength", values_to = "absorbance") %>%
  dplyr::mutate(wavelength = gsub("scan_nir.|_abs", "", wavelength)) %>%
  dplyr::mutate(wavelength = as.numeric(wavelength)) %>%
  ggplot(aes(x = wavelength, y = absorbance, group = id.layer_local_c)) +
  geom_line(alpha = 0.1, color = "darkblue") +
  scale_x_continuous(breaks = seq(680, 2500, by = 200))+
  labs(title = "NIR Spectra (100 random scans)",
       x = "Wavelength (nm)",
       y = "Absorbance")+
  theme_light()
```


```{r}
toc()
rm(list = ls())
gc()
```

## References


